---
title: 'WP1: Simple View - Latent Profile Analysis'
output: 
  html_document:
    toc: true
    toc_float: true
    
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output/analysis/") })
---

This script is used for running the initial latent profile models. It will only be used *if* the CFA indicates that a single factor is best, otherwise we will use a factor mixture model (WP1_SimpView_FMM1.Rmd).

These models will be run twice: once using the ML estimator and once using the MLR estimator (changed in the [lpa-base chunk](#base)). This will allow us to assess for departure from multivariate normality in the next stage of model fitting (WP1_SimpView_LPA2.Rmd). 

# Set-up

```{r libraries}
# Specify packages
pkgs <- c("tidyverse", "MplusAutomation", "texreg", "DiagrammeR", "kableExtra", "ggalluvial", "scales")

# Use groundhog package to load correct package versions
# if (!require("groundhog")) install.packages("groundhog")
# groundhog.day <- "2021-03-01"

# groundhog::groundhog.library(pkgs, groundhog.day)

# Or if not using groundhog package - load currently installed package versions separately (reproducibility not guaranteed)
invisible(lapply(pkgs, library, character.only = TRUE))
```

```{r create-dir}
# Create subdirectories for storing mplus scripts, data files, and output, if do not already exist
if(dir.exists("./mplus_models/")==FALSE){dir.create("./mplus_models/")}
if(dir.exists("./mplus_models/simpview/")==FALSE){dir.create("./mplus_models/simpview/")}
if(dir.exists("./mplus_models/simpview/lpa")==FALSE){dir.create("./mplus_models/simpview/lpa")} 
if(dir.exists("../output/")==FALSE){dir.create("../output/")}
if(dir.exists("../output/figures")==FALSE){dir.create("../output/figures")}
if(dir.exists("../output/tables")==FALSE){dir.create("../output/tables")}
```

###  Load and prepare data for Mplus

```{r load-data}
sv_data <- read.csv("../data/simulated/WP1_SimpView_sim_raw_n1000_k3.csv")
# sv_data <- read.csv("../data/processed/WP1_data.csv") %>% 
#  select(yp_id, yp_no, age_m_f8, age_m_f9, nara_comp_raw_f9, nara_acc_raw_f9, read_word_raw_f9, read_nonw_raw_f9, wold_comp_raw_f8)
```

Mplus has an 8-character limit, so rename relevant variables with short names for modelling. 

```{r mplus-names}
# Check names
str_sub(names(sv_data), 1, 8)

# Rename any necessary, print new names
sv_data <- sv_data %>% 
  rename(naraComp = nara_comp_raw_f9,
         naraAcc = nara_acc_raw_f9,
         wordAcc = read_word_raw_f9,
         nonwAcc = read_nonw_raw_f9,
         woldComp = wold_comp_raw_f8,
         f8age = age_m_f8,
         f9age = age_m_f9) 

names(sv_data)
```

Variable format problematic for mplus data conversion, change ID to factor.

```{r mplus-formatting}
sv_data <- sv_data %>% 
  mutate(yp_id = as.factor(yp_id))
```


# Latent Profile Model to extract latent classes

**Aim:** to fit models of different n classes, at range of different model specifications.

Indicator variances can always differ from each other within a class, but different model specifications vary in whether they are separately estimated between classes. Labels for different specifications taken from Pastor (2007) and Masyn (2013).

### Base model {#base}

Set up initial model, based on CFA above. 

*Note:* Increase starts to 2000 200 for a model if best log-likelihood value not replicated (Lubke & Luningham, 2017)

```{r lpa-base}
# Settings 
starts = "500 50"   # Increase if necessary
estimator = "ml"   # Run with ml and mlr

# Create base model
m_lpa_base <- mplusObject(
  TITLE = "Latent Profile Analysis;",
  ANALYSIS = sprintf("estimator = %s; type = mixture; starts = %s;", estimator, starts),
  VARIABLE = "classes = c(1);",
  MODEL = "%OVERALL% 
           wordAcc nonwAcc naraAcc naraComp on f9age;
           woldcomp on f8age;", 
  OUTPUT = "stdyx; residual; TECH1; TECH7; TECH8; TECH11;",
  PLOT = "TYPE = PLOT3;",
  usevariables = colnames(sv_data[,!names(sv_data) %in% c("yp_id", "yp_no", "k")]),
  rdata = sv_data)

# Specify all variances
lpa_vars <- "naraComp; naraAcc; wordAcc; nonwAcc; woldComp;"

# Specify all covariances 
lpa_covars <- "
  naraComp with naraAcc;
  naraComp with wordAcc; 
  naraComp with nonwAcc; 
  naraComp with woldComp; 
  naraAcc with wordAcc; 
  naraAcc with nonwAcc; 
  naraAcc with woldComp; 
  wordAcc with nonWacc; 
  wordAcc with woldComp; 
  nonwAcc with woldComp; "
```

Follow class-enumeration process set out by Masyn (2013). Use a one-class LPA (incorporating sample means and covariances) as an absolute fit benchmark.

```{r lpa-benchmark}
m_lpa_benchmark <- update(m_lpa_base,
     TITLE = ~ "LPA Benchmark",
     VARIABLE = ~ "classes = c(1);",
     MODEL = as.formula(sprintf("~ . + '%%OVERALL%% \n %s \n %s'", lpa_vars, lpa_covars)))

mplusModeler(m_lpa_benchmark, paste0("./mplus_models/simpview/lpa/sv_lpa_", estimator, "_mbenchmark.dat"), run = TRUE)
```

### Model A: Class-invariant, diagonal

Indicator variances differ from each other within a class, but are constrained to be equal across classes. No covariances are estimated. 

```{r lpa-a-fit}
# Fit models
m_lpa_a <- lapply(1:6, function(k) {
   body <- update(m_lpa_base,
     TITLE = as.formula(sprintf("~ 'LPA Model A: %d classes;'", k)),
     VARIABLE = as.formula(sprintf("~ 'classes = c(%d);'", k)))
   
   mplusModeler(body, sprintf("mplus_models/simpview/lpa/sv_lpa_%s_a_%dclass.dat", estimator, k), run = TRUE)
 })

```

```{r lpa-a-check}
# Read in output
lpa_a_out <- readModels(target = "./mplus_models/simpview/lpa", filefilter = sprintf("sv_lpa_%s_a", estimator))

# Check warnings
for (model in 1:length(lpa_a_out)){
  print(lpa_a_out[[model]]$input$title)
  print(lpa_a_out[[model]]$warnings)
}
```

### LPA Model B: Class-invariant, unrestricted

Indicator variances differ from each other within a class and can covary, but variances and covariances constrained to be equal across classes.

```{r lpa-b-fit}
m_lpa_b <- lapply(1:6, function(k) {
   body <- update(m_lpa_base,
     TITLE = as.formula(sprintf("~ 'LPA Model B: %d classes;'", k)),
     VARIABLE = as.formula(sprintf("~ 'classes = c(%d);'", k)),
     
     # Update model spec 1: overall model estimates (variances and covariances)
     MODEL = as.formula(sprintf("~ . + '\n %s \n %s'", lpa_vars, lpa_covars)))
   
   # Run model
   mplusModeler(body, sprintf("mplus_models/simpview/lpa/sv_lpa_%s_b_%dclass.dat", estimator, k), run = TRUE)
 })
```

```{r lpa-b-check}
# Read in output
lpa_b_out <- readModels(target = "./mplus_models/simpview/lpa", filefilter = sprintf("sv_lpa_%s_b", estimator))

# Check warnings
for (model in 1:length(lpa_b_out)){
  print(lpa_b_out[[model]]$input$title)
  print(lpa_b_out[[model]]$warnings)
}
```

### LPA Model C: Class-varying, diagonal 

Indicator variances are estimated separately for each class. No covariances are estimated.

```{r lpa-c-fit}
m_lpa_c <- lapply(1:6, function(k) {
  
  # Update model spec 1: overall model estimates (variances only)
  body <- update(m_lpa_base,
              MODEL = as.formula(sprintf("~ . + '%s'", lpa_vars)))
  
  # Create class-level specifications (to include variances only)
  class_spec <- paste0("%c#1% \n ", lpa_vars)
  if (k > 1){
    for (i in 2:k){
      class_spec <- paste0(class_spec, "\n %c#", i, "% \n ", lpa_vars)
    }
  }
  
  # ...Update model spec 2: class estimates
  body <- update(body,
     TITLE = as.formula(sprintf("~ 'LPA Model C: %d classes;'", k)),
     VARIABLE = as.formula(sprintf("~ 'classes = c(%d);'", k)),
     MODEL = as.formula(sprintf("~ . + '%s'", class_spec)))
   
  # Run model
  mplusModeler(body, sprintf("mplus_models/simpview/lpa/sv_lpa_%s_c_%dclass.dat", estimator, k), run = TRUE)
 })

```

```{r lpa-c-check}
# Read in output
lpa_c_out <- readModels(target = "./mplus_models/simpview/lpa", filefilter = sprintf("sv_lpa_%s_c", estimator))

# Check warnings
for (model in 1:length(lpa_c_out)){
  print(lpa_c_out[[model]]$input$title)
  print(lpa_c_out[[model]]$warnings)
}
```

### LPA Model D 

Indicator variances are estimated separately for each class, but covariances are constrained to be equal across classes.

```{r lpa-d-fit}
m_lpa_d <- lapply(1:6, function(k) {
  
  # Update model spec 1: overall model estimates (variances and covariances)
  body <- update(m_lpa_base,
              MODEL = as.formula(sprintf("~ . + '%s \n %s '", lpa_vars, lpa_covars)))   
  
  # Create class-level specifications (variances only)
  class_spec <- paste0("%c#1% \n ", lpa_vars)
  if (k > 1){
    for (i in 2:k){
      class_spec <- paste0(class_spec, "\n %c#", i, "% \n ", lpa_vars)
    }
  }
  
  # ...Update model spec 2: class estimates
  body <- update(body,
     TITLE = as.formula(sprintf("~ 'LPA Model D: %d classes;'", k)),
     VARIABLE = as.formula(sprintf("~ 'classes = c(%d);'", k)),
     MODEL = as.formula(sprintf("~ . + '%s'", class_spec)))
  
  # Run model
  mplusModeler(body, sprintf("mplus_models/simpview/lpa/sv_lpa_%s_d_%dclass.dat", estimator, k), run = TRUE)
 })
```

```{r lpa-d-check}
# Read in output
lpa_d_out <- readModels(target = "./mplus_models/simpview/lpa", filefilter = sprintf("sv_lpa_%s_d", estimator))

# Check warnings
for (model in 1:length(lpa_d_out)){
  print(lpa_d_out[[model]]$input$title)
  print(lpa_d_out[[model]]$warnings)
}
```


### LPA Model E: Class-varying, unrestricted

All variances and covariances can vary between clusters. 

```{r lpa-e-fit}
m_lpa_e <- lapply(1:6, function(k) {
  
  # Update model spec 1: overall model estimates (variances and covariances)
  body <- update(m_lpa_base,
              MODEL = as.formula(sprintf("~ . + '%s \n %s '", lpa_vars, lpa_covars)))   
  
  # Create class-level specifications (variances and covariances)
  class_spec <- paste0("%c#1% \n ", lpa_vars,  lpa_covars)
  if (k > 1){
    for (i in 2:k){
      class_spec <- paste0(class_spec, "\n %c#", i, "% \n ", lpa_vars, lpa_covars)
    }
  }
  
  # ...Update model spec 2: class estimates
  body <- update(body,
     TITLE = as.formula(sprintf("~ 'LPA Model E: %d classes;'", k)),
     VARIABLE = as.formula(sprintf("~ 'classes = c(%d);'", k)),
     MODEL = as.formula(sprintf("~ . + '%s'", class_spec)))

  # Run model
  mplusModeler(body, sprintf("mplus_models/simpview/lpa/sv_lpa_%s_e_%dclass.dat", estimator, k), run = TRUE)
 })
```

```{r lpa-e-check}
# Read in output
lpa_e_out <- readModels(target = "./mplus_models/simpview/lpa", filefilter = sprintf("sv_lpa_%s_e", estimator))

# Check warnings
for (model in 1:length(lpa_e_out)){
  print(lpa_e_out[[model]]$input$title)
  print(lpa_e_out[[model]]$warnings)
}
```

