---
title: "WP1: Profiles of comprehension difficulty - Data Pre-processing and CFA"
output: 
  html_document:
    toc: true
    toc_float: true
    
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output/analysis/") })
---

This script computes basic summary statistics for the data, ensures that they align with Mplus output, and tests the factor structure to be used for subsequent analyses. 

# Set-up

```{r libraries}
# List packages
pkgs <- c("tidyverse", "psych", "naniar", "MplusAutomation", "texreg", "semPlot")

# Use groundhog package to load correct package versions
# if (!require("groundhog")) install.packages("groundhog")
# groundhog.day <- "2021-03-01"
# groundhog::groundhog.library(pkgs, groundhog.day)

# Or if not using groundhog package - load currently installed package versions separately (reproducibility not guaranteed)
invisible(lapply(pkgs, library, character.only = TRUE))
```

```{r create-dir}
# Create subdirectories for storing mplus scripts, data files, and output, if do not already exist
if(dir.exists("./mplus_models/")==FALSE){dir.create("./mplus_models/")}
if(dir.exists("./mplus_models/desc")==FALSE){dir.create("./mplus_models/desc")}
if(dir.exists("./mplus_models/cfa")==FALSE){dir.create("./mplus_models/cfa")} 
```

###  Load and prepare data for Mplus

```{r load-data}
cp_data <- read.csv("../data/simulated/WP1_CompProf_sim_raw_n694_k3.csv")
#cp_data <- read.csv("../data/processed/WP1_data.csv")
```

Mplus has an 8-character limit, so rename relevant variables with short names for modelling. 

```{r mplus-names}
# Rename any necessary, print new names
cp_data <- cp_data %>% 
  rename(naraComp = nara_comp_raw_f9,
         naraAcc = nara_acc_raw_f9,
         wordAcc = read_word_raw_f9,
         nonwAcc = read_nonw_raw_f9,
         woldComp = wold_comp_raw_f8,
         f8age = age_m_f8,
         f9age = age_m_f9,
         f10age = age_m_f10,
         naraRate = nara_rate_raw_f9,
         woldVcb = wold_vcb_raw_f8,
         wiscVcb = wisc_vcb_raw_f8,
         wiscPcmp = wisc_pcmp_raw_f8,
         wiscCode = wisc_code_raw_f8,
         wiscParr = wisc_parr_raw_f8,
         wiscBloc = wisc_bloc_raw_f8,
         wiscObja = wisc_obja_raw_f8,
         wiscBwsp = wisc_bwsp_raw_f8,
         cntspan = cntsp_span_raw_f10,
         attnSel = teach_slct_raw_f8,
         attnDiv = teach_divd_raw_f8,
         attnOpp = teach_ctr_diff_f8,
         sdqHyp = sdq_hyp_prnt_ku) 

# Check compatibility with mplus
str_sub(names(cp_data), 1, 8)
```

Variable format problematic for mplus data conversion, change ID to factor.

```{r mplus-formatting}
cp_data <- cp_data %>% 
  mutate(yp_id = as.factor(yp_id))
```


# Descriptive statistics

Extract summary statistics for each variable.

```{r summary-stats}
describe(cp_data)
vis_miss(cp_data, cluster = TRUE)
```

Patterns of missingness are more thoroughly explored for all WP1 variables in the initial data processing script. We anticipate more missingness than in the SimpView analysis as there were more measures collected from across three different clinics. 

### Check that mplus reading data as expected

Extract descriptive statistics as check.

```{r mplus-desc}
# Specify model
m_desc <- mplusObject(
  TITLE = "Data check - Descriptive statistics;",
  ANALYSIS = "type = basic;",
  usevariables = colnames(cp_data[,!names(cp_data) %in% c("yp_id", "yp_no", "k")]),
  rdata = cp_data)

# Fit model 
m_desc_fit <- mplusModeler(m_desc,
                            dataout = "./mplus_models/desc/cp_sim.dat",
                            modelout = "./mplus_models/desc/cp_check.inp",
                            check = TRUE, run = TRUE, hashfilename = TRUE)

# Read mplus output
m_desc_out <- readModels("./mplus_models/desc/cp_check.out")

# Check that sample sizes and means match descriptive statistics from above
r_summ <- describe(cp_data) %>%
  as.data.frame() %>% 
  slice(3:26) %>% 
  select(n, mean) %>% 
  round(., 2)
mplus_summ <- m_desc_out$sampstat$univariate.sample.statistics %>% 
  as.data.frame() %>%
  round(.,2) %>% 
  select(`Sample Size`, Mean) %>% 
  set_names(c("n", "mean"))

if (any((r_summ == mplus_summ) == FALSE)){
  print("WARNING: DIFFERENCES DETECTED BETWEEN R AND MPLUS DATA SUMMARIES. Inspect output.")
  setdiff(mplus_summ, r_summ)
  } else{
      print("Data check passed")
  }
```

# Confirmatory Factor Analysis

### Testing the factor structure

Test whether the data support the proposed factorial model over a single reading factor.

A good-fitting model should ideally meet the following criteria:
* CFI & TLI > .95  
* RMSEA <= .05  
* SRMR < .05

```{r cfa-factor-structure}
# ONE-FACTOR MODEL
m_cfa1 <- mplusObject(
  TITLE = "Confirmatory Factor Analysis - Single Factor;",
  ANALYSIS = "estimator = mlr; type = general;",
  MODEL = "wordAcc nonwAcc naraAcc naraComp naraRate on f9age;
           woldComp woldVcb wiscVcb wiscPcmp wiscCode wiscParr 
           wiscBloc wiscObja wiscBwsp attnSel attnDiv attnOpp on f8age;
           cntSpan on f10age;
           read by naraComp naraAcc wordAcc nonwAcc woldComp naraRate 
           woldVcb wiscVcb wiscPcmp wiscCode wiscParr wiscBloc wiscObja 
           wiscBwsp cntSpan attnSel attnDiv attnOpp sdqHyp;",
  OUTPUT = "sampstat; TECH1; TECH4; stdyx; modindices; ",
  PLOT = "TYPE = PLOT3;",
  usevariables = colnames(cp_data[,!names(cp_data) %in% c("yp_id", "yp_no", "k")]),
  rdata = cp_data)

m_cfa1_fit <- mplusModeler(m_cfa1,
                           modelout = "./mplus_models/cfa/cp_cfa1.inp",
                           check = TRUE, run = TRUE)

# TWO-FACTOR MODEL
m_cfa2 <- update(m_cfa1,
                 TITLE = ~ "Confirmatory Factor Analysis - Six-Factor;",
                 ANALYSIS = ~ . + "starts = 20;",
                 MODEL = ~ "wordAcc nonwAcc naraAcc naraComp naraRate on f9age;
                            woldComp woldVcb wiscVcb wiscPcmp wiscCode wiscParr 
                            wiscBloc wiscObja wiscBwsp attnSel attnDiv attnOpp on f8age;
                            cntSpan on f10age;
                            acc by nonwAcc wordAcc naraAcc; 
                            comp by naraComp woldComp;
                            naraRate;
                            vocab by wiscVcb woldVcb ;
                            perfiq by wiscPcmp wiscCode wiscParr wiscBloc wiscObja;
                            execfun by wiscBwsp cntSpan attnSel attnDiv attnOpp sdqHyp;
                            acc with comp; acc with naraRate; acc with vocab; 
                            acc with perfiq; acc with execfun;
                            comp with naraRate; comp with vocab; 
                            comp with perfiq; comp with execfun;
                            naraRate with vocab; naraRate with perfiq; 
                            naraRate with execfun; 
                            vocab with perfiq; vocab with execfun;
                            perfiq with execfun;") # 6-factor
m_cfa2_fit <- mplusModeler(m_cfa2,
                           modelout = "./mplus_models/cfa/cp_cfa2.inp",
                           check = TRUE, run = TRUE)

# COMPARE MODELS
cfa_models <- readModels(target = "./mplus_models/cfa", filefilter = "cp_cfa")
SummaryTable(cfa_models, keepCols = c("Title", "Parameters", "LL", "CFI", "TLI", "AIC", "BIC", "RMSEA_Estimate", "RMSEA_pLT05", "SRMR"))
```

* Remove indicators that do not load onto their corresponding factors (0.30 or 0.32 cut-off??)
* >> If multiple, can trial adding in as separate factor if theoretically sensible
* If not key indicator of interest (e.g., TEACh scales) and do not fit well >> discard???


If a one-factor model is preferred, then we will proceed with using a Latent Profile Analysis. If the data support a two-factor model, as anticipated, then we will use a Factor Mixture Model. Consider parameters that might improve model fit for the FMM. 

### Improving fit 

```{r cfa-inspection}
# Inspect model (by = loadings, with = covariances)
cfa_models$cp_cfa2.out$parameters$unstandardized
cfa_models$cp_cfa2.out$parameters$stdyx.standardized

# Inspect modification indices
cfa_models$cp_cfa2.out$mod_indices %>% 
  filter(operator == "WITH") %>% 
  arrange(desc(MI))
```
* Check for very large standardised residuals
* Use MIs to change only one at a time, starting with largest, but must be theoretically meaningful. 

E.g...

```{r cfa-modification1}
# TWO-FACTOR MODEL with additional covariance between the two working memory measures
m_cfa2b <- update(m_cfa2,
                 TITLE = ~ "Confirmatory Factor Analysis - Two-Factor - modification1 working memory;",
                 MODEL = ~. + "cntSpan with wiscbwsp;")

m_cfa2b_fit <- mplusModeler(m_cfa2b,
                           modelout = "./mplus_models/cfa/cp_cfa2b.inp",
                           check = TRUE, run = TRUE)

# COMPARE MODELS
cfa_models <- readModels(target = "./mplus_models/cfa", filefilter = "cp_cfa")
SummaryTable(cfa_models, keepCols = c("Title", "Parameters", "LL", "CFI", "TLI", "AIC", "BIC", "RMSEA_Estimate", "RMSEA_pLT05", "SRMR"))

# Inspect model (by = loadings, with = covariances)
cfa_models$cp_cfa2b.out$parameters$unstandardized
cfa_models$cp_cfa2b.out$parameters$stdyx.standardized

# Inspect modification indices
cfa_models$cp_cfa2b.out$mod_indices %>% 
  filter(operator == "WITH") %>% 
  arrange(desc(MI))
```
```{r cfa-modification2}
# TWO-FACTOR MODEL with additional covariance between the two working memory measures
m_cfa2c <- update(m_cfa2b,
                 TITLE = ~ "Confirmatory Factor Analysis - Two-Factor - modification2 NARA rate-acc;",
                 MODEL = ~. + "naraRate with naraAcc;")

m_cfa2c_fit <- mplusModeler(m_cfa2c,
                           modelout = "./mplus_models/cfa/cp_cfa2c.inp",
                           check = TRUE, run = TRUE)

# COMPARE MODELS
cfa_models <- readModels(target = "./mplus_models/cfa", filefilter = "cp_cfa")
SummaryTable(cfa_models, keepCols = c("Title", "Parameters", "LL", "CFI", "TLI", "AIC", "BIC", "RMSEA_Estimate", "RMSEA_pLT05", "SRMR"))

# Inspect model (by = loadings, with = covariances)
cfa_models$cp_cfa2c.out$parameters$unstandardized
cfa_models$cp_cfa2c.out$parameters$stdyx.standardized

# Inspect modification indices
cfa_models$cp_cfa2c.out$mod_indices %>% 
  filter(operator == "WITH") %>% 
  arrange(desc(MI))
```


A good-fitting model should ideally meet the following criteria:
* CFI & TLI > .95  
* RMSEA <= .05  
* SRMR < .05

### CFA figure

```{r cfa-diagram}
# semPaths(semPlotModel(cfa_models$cp_cfa1.out, mplusStd = "stdyx"), what = "paths", whatLabels = "std", rotation = 2, intercepts = FALSE)
```

