---
title: "WP1: Simple View - Data Pre-processing and CFA"
output: 
  html_document:
    toc: true
    toc_float: true
date: '06/05/2021'
      
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output/analysis/") })
---

This script computes basic summary statistics for the data, ensures that they align with Mplus output, and tests the factor structure to be used for subsequent analyses. This analysis is conducted using the "exploratory" or "calibration" half of the dataset (subsample A).

# Set-up

```{r libraries}
# List packages
pkgs <- c("tidyverse", "psych", "naniar", "MplusAutomation", "texreg", "semPlot")

# Use groundhog package to load correct package versions
if (!require("groundhog")) install.packages("groundhog")
groundhog.day <- "2021-03-01"
groundhog::groundhog.library(pkgs, groundhog.day)

# Or if not using groundhog package - load currently installed package versions separately (reproducibility not guaranteed)
# invisible(lapply(pkgs, library, character.only = TRUE))
```

```{r create-dir}
# Create subdirectories for storing mplus scripts, data files, and output, if do not already exist
if(dir.exists("./mplus_models/")==FALSE){dir.create("./mplus_models/")}
if(dir.exists("./mplus_models/simpview/")==FALSE){dir.create("./mplus_models/simpview/")}
if(dir.exists("./mplus_models/simpview/desc")==FALSE){dir.create("./mplus_models/simpview/desc")}
if(dir.exists("./mplus_models/simpview/cfa")==FALSE){dir.create("./mplus_models/simpview/cfa")} 
```

###  Load and prepare data for Mplus

```{r load-data}
#sv_data <- read.csv("../data/simulated/WP1_SimpView_sim_raw_n1000_k3.csv") # for script testing 
#sv_data <- read.csv("../data/processed/WP1_data_subA.csv")                 # ALSPAC data
```

Mplus has an 8-character limit, so rename relevant variables with short names for modelling. 

```{r mplus-names}
# Check names
str_sub(names(sv_data), 1, 8)

# Rename any necessary, print new names
sv_data <- sv_data %>% 
  rename(naraComp = nara_comp_raw_f9,
         naraAcc = nara_acc_raw_f9,
         wordAcc = read_word_raw_f9,
         nonwAcc = read_nonw_raw_f9,
         woldComp = wold_comp_raw_f8,
         f8age = age_m_f8,
         f9age = age_m_f9) 

names(sv_data)
```

Variable format problematic for mplus data conversion, change ID to factor.

```{r mplus-formatting}
sv_data <- sv_data %>% 
  mutate(yp_id = as.factor(yp_id))
```


# Descriptive statistics

Extract summary statistics for each variable.

```{r summary-stats}
describe(sv_data)
vis_miss(sv_data)
```

Patterns of missingness are more thoroughly explored for all WP1 variables in the initial data processing script. For this particular analysis, we anticipate relatively low levels of missingness as measures were collected close in time (all at age 9 clinic, except WOLD comprehension at age 8 clinic), and that this will likely be MCAR. Missingness will be dealt with using full information maximum likelihood. 

### Check that mplus reading data as expected

Extract descriptive statistics as check.

```{r mplus-desc}
# Specify model
m_desc <- mplusObject(
  TITLE = "Data check - Descriptive statistics;",
  ANALYSIS = "type = basic;",
  usevariables = c("naraComp", "naraAcc", "wordAcc", "nonwAcc", "woldComp", "f8age", "f9age"),
  rdata = sv_data)

# Fit model 
m_desc_fit <- mplusModeler(m_desc,
                            dataout = "./mplus_models/simpview/desc/sv_sim.dat",
                            modelout = "./mplus_models/simpview/desc/sv_check.inp",
                            check = TRUE, run = TRUE, hashfilename = TRUE)

# Read mplus output
m_desc_out <- readModels("./mplus_models/simpview/desc/sv_check.out")

# Check that sample sizes and means match descriptive statistics from above
r_summ <- describe(sv_data) %>%
  as.data.frame() %>% 
  slice(3:9) %>% 
  select(n, mean) %>% 
  round(., 2)
mplus_summ <- m_desc_out$sampstat$univariate.sample.statistics %>% 
  as.data.frame() %>%
  round(.,2) %>% 
  select(`Sample Size`, Mean)

if (any((r_summ - mplus_summ) > abs(0.01))){  # allowing for rounding errors
  print("WARNING: DIFFERENCES DETECTED BETWEEN R AND MPLUS DATA SUMMARIES")
  } else{
      print("Data check passed")
  }
```

# Confirmatory Factor Analysis

### One vs. Two Factors

Test whether the data support the proposed two-factor model (accuracy, comprehension) over a single reading factor.
Each model additionally includes age as a covariate.

```{r cfa-factor-structure}
# ONE-FACTOR MODEL
m_cfa1 <- mplusObject(
  TITLE = "Confirmatory Factor Analysis - Single Factor;",
  ANALYSIS = "estimator = mlr; type = general;",                  
  MODEL = "read by naraComp naraAcc wordAcc nonwAcc woldComp;
           wordAcc nonwAcc naraAcc naraComp on f9age;
           woldcomp on f8age;",   # single reading factor, plus age covariates 
  OUTPUT = "sampstat; TECH1; TECH4; stdyx; modindices; ",
  PLOT = "TYPE = PLOT3;",
  usevariables = colnames(sv_data[,!names(sv_data) %in% c("yp_id", "yp_no")]),
  rdata = sv_data)

m_cfa1_fit <- mplusModeler(m_cfa1,
                           modelout = "./mplus_models/simpview/cfa/sv_cfa1.inp",
                           check = TRUE, run = TRUE)

# TWO-FACTOR MODEL
m_cfa2 <- update(m_cfa1,
                 TITLE = ~ "Confirmatory Factor Analysis - Two-Factor;",
                 MODEL = ~ "acc by nonwAcc wordAcc naraAcc; comp by woldComp naraComp;
                            wordAcc nonwAcc naraAcc naraComp on f9age;
                            woldcomp on f8age;
                            acc with comp;") # two factors, plus age covariates
m_cfa2_fit <- mplusModeler(m_cfa2,
                           modelout = "./mplus_models/simpview/cfa/sv_cfa2.inp",
                           check = TRUE, run = TRUE)

# COMPARE MODELS
cfa_models <- readModels(target = "./mplus_models/simpview/cfa", filefilter = "sv_cfa")
SummaryTable(cfa_models, keepCols = c("Title", "Parameters", "LL", "CFI", "TLI", "AIC", "BIC", "RMSEA_Estimate", "RMSEA_pLT05", "SRMR"))
```

If a one-factor model is preferred, then we will proceed with using a Latent Profile Analysis. If the data support a two-factor model, as anticipated, then we will use a Factor Mixture Model. If proceeding with FMM, consider parameters that might improve model fit.

### Improving fit for a Two-Factor Model

```{r cfa-inspection}
# Inspect model (by = loadings, with = covariances)
cfa_models$sv_cfa2.out$parameters$unstandardized
cfa_models$sv_cfa2.out$parameters$stdyx.standardized

# Inspect modification indices
cfa_models$sv_cfa2.out$mod_indices
```

* Check for very large standardised residuals.  
* No accepted view on how large MIs should be. Change only one at a time, starting with largest, but must be theoretically meaningful.  
* *A priori*, we consider possible meaningful influences to be to be (in this order): naracomp with naraacc (taken from same assessment), wordacc with nonwacc (same battery of tests, similar format), wordacc with naraacc (both real word accuracy measures).  

E.g...

```{r cfa-modification}
# TWO-FACTOR MODEL with additional covariance between NARA accuracy and comprehension
# m_cfa2b <- update(m_cfa2,
#                  TITLE = ~ "Confirmatory Factor Analysis - Two-Factor - modification1 nara covariance;",
#                  MODEL = ~. + "naraAcc with naraComp;")
# 
# m_cfa2b_fit <- mplusModeler(m_cfa2b,
#                            modelout = "./mplus_models/simpview/cfa/sv_cfa2b.inp",
#                            check = TRUE, run = TRUE)
# 
# # COMPARE MODELS
# cfa_models <- readModels(target = "./mplus_models/simpview/cfa", filefilter = "sv_cfa")
# SummaryTable(cfa_models, keepCols = c("Title", "Parameters", "LL", "CFI", "TLI", "AIC", "BIC", "RMSEA_Estimate", "RMSEA_pLT05", "SRMR"))
# 
# # Inspect model (by = loadings, with = covariances)
# cfa_models$sv_cfa2b.out$parameters$unstandardized
# cfa_models$sv_cfa2b.out$parameters$stdyx.standardized
# 
# # Inspect modification indices
# cfa_models$sv_cfa2b.out$mod_indices
```

The following rules of thumb are often used to indicate a good-fitting model:
* CFI & TLI > .95  
* RMSEA <= .05  
* SRMR < .05

### CFA figure

```{r cfa-diagram}
# semPaths(semPlotModel(cfa_models$sv_cfa2b.out, mplusStd = "stdyx"), what = "paths", whatLabels = "std", rotation = 1, intercepts = FALSE)
```

