---
title: 'WP1: Simple View - Final Class Model'
output: 
  html_document:
    toc: true
    toc_float: true
    
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output/analysis/") })
---

This script is used for extracting descriptive statistics for the final class model (LPA or FMM).

# Set-up

```{r libraries}
# Specify packages
pkgs <- c("tidyverse", "MplusAutomation", "texreg", "kableExtra")

# Use groundhog package to load correct package versions
# if (!require("groundhog")) install.packages("groundhog")
# groundhog.day <- "2021-03-01"
# groundhog::groundhog.library(pkgs, groundhog.day)

# Or if not using groundhog package - load currently installed package versions separately (reproducibility not guaranteed)
invisible(lapply(pkgs, library, character.only = TRUE))
```

Read in selected model file that contains class assignment data, and data file of descriptive statistics. Merge and check. 

```{r read-data}
# Read in and extract class data
final_mod <- readModels(target = "./mplus_models/simpview/shape", filefilter = "lpa4_rerun_5")  # final model file
class_data <- as.data.frame(final_mod$savedata) %>% 
  rename_all(tolower) %>% 
  select(cidb3153, naraacc:c, f8age, f9age) %>%   # keep age data for additional checking during merge
  rename(cidB3153 = cidb3153)

# Read in and extract relevant variables from data file
ppt_data <- read.csv("../data/processed/WP1_data_subA.csv") %>%
  select(cidB3153, age_m_f8, age_m_f9, nara_comp_raw_f9, nara_acc_raw_f9, read_word_raw_f9, read_nonw_raw_f9, wold_comp_raw_f8,
         nara_comp_std_f9, nara_acc_std_f9) %>% 
  mutate(read_comb_raw_f9 = (read_word_raw_f9 + read_nonw_raw_f9))

### ... simulated version for testing
# ppt_data <- read.csv("../data/simulated/WP1_SimpView_sim_raw_n1000_k3.csv") %>%
#   select(yp_id, yp_no, age_m_f8, age_m_f9, nara_comp_raw_f9, nara_acc_raw_f9, read_word_raw_f9, read_nonw_raw_f9, wold_comp_raw_f8) %>% 
#   mutate(nara_comp_std_f9 = round(rnorm(n(), 100, 15)), # Create dummy variables for now
#          nara_acc_std_f9 = round(rnorm(n(), 100, 15)))  

# Merge
ppt_classes <- ppt_data %>% 
  full_join(class_data, by = c("cidB3153",
                             "age_m_f8" = "f8age",
                             "age_m_f9" = "f9age"))
```


# Whole sample descriptives

```{r all-desc}
ppt_classes %>% 
  select(-cidB3153, -c(cprob1:c)) %>% 
  psych::describe() %>% 
  as.data.frame() %>% 
  select(n, mean, sd, min, max) %>% 
  round(., 2)
```

The NARA comes with test-normed standardised scores, but the others have raw score only (bespoke tests and/or only subset of items). Compute sample standardised scores, accounting for age variability.

```{r standardise}
sample_std <- function(test_scores = NA, age = NA, data = NA){
  
  # Create row references
  data <- data %>% 
    rownames_to_column("rowno_id")
  
  # For each variable in list
  for (test in test_scores){
    
      # Create new variable name
      var_name = str_replace(test, "raw", "bstd")   # for bespoke standard
    
      # Regress out age
      lm_age <- lm(get(test) ~ get(age), data = data)
      
      # Use standardised residuals, and then convert to standardised score (M = 100, SD = 15)
      ppt_resid <- rstandard(lm_age)
      std_score <- round(ppt_resid*15 + 100)

      # Convert new standardised score to data frame
      append_score <- data.frame(std_score) %>%
        set_names(var_name) %>%
        rownames_to_column("rowno_id")

      # Merge score by row reference
      data <- data %>%
        left_join(append_score, by = "rowno_id") 
  }
  
  # Remove row reference from final output
  data <- data %>% 
    select(-rowno_id)
}

std_data <- sample_std(data = ppt_classes, test_scores = c("read_comb_raw_f9", "nara_acc_raw_f9", "nara_comp_raw_f9"), age = "age_m_f9")
std_data <- sample_std(data = std_data, test_scores = c("wold_comp_raw_f8"), age = "age_m_f8")
```


Intend to report test-standardised scores for the NARA, but check how the two compare. 
```{r std-compare}
# Accuracy
plot(std_data$nara_acc_std_f9, std_data$nara_acc_bstd_f9) +
  abline(1, 1)

# Comprehension
plot(std_data$nara_comp_std_f9, std_data$nara_comp_bstd_f9) +
  abline(1, 1)
```

# Class descriptives 

```{r class-desc}
class_averages_raw <- ppt_classes %>% 
  group_by(c) %>% 
  summarise_at(vars(read_nonw_raw_f9, read_word_raw_f9, nara_acc_raw_f9, nara_comp_raw_f9, wold_comp_raw_f8 ), 
               list(mean = mean, sd = sd, min = min, max = max), na.rm = TRUE) %>% 
  round(.,2)

class_averages_raw

class_averages_std <- std_data %>% 
  group_by(c) %>% 
  summarise_at(vars(read_comb_bstd_f9, nara_acc_bstd_f9, nara_comp_bstd_f9, wold_comp_bstd_f8), 
               list(mean = mean, sd = sd, min = min, max = max), na.rm = TRUE) %>% 
  round(.,2)

class_averages_std
```

```{r class-plot}
#### STILL TO-DO / TO MAKE NICE ####
# would be nice to add distributions, but might end up looking messy

# Format data for plotting
plot_data <- std_data %>% 
  select(cidB3153, c, read_comb_bstd_f9, nara_acc_bstd_f9, nara_comp_bstd_f9, wold_comp_bstd_f8) %>% 
  pivot_longer(cols = c(read_comb_bstd_f9,  nara_acc_bstd_f9, nara_comp_bstd_f9, wold_comp_bstd_f8),
               names_to = "task",
               values_to = "score") %>% 
  group_by(c, task) %>%
  summarise(class_mean = mean(score, na.rm = TRUE), sd = sd(score, na.rm = TRUE), se = sd/sqrt(n()),
            n = n(), proportion = n/3423*100) %>% 
  mutate(c = paste0(c, " (", round(proportion,2), "%)")) %>% 
  mutate(c = as.factor(c),
         task = factor(task, levels = c("read_comb_bstd_f9", "nara_acc_bstd_f9", "nara_comp_bstd_f9", "wold_comp_bstd_f8")))

long_data <- std_data %>% 
  select(cidB3153, c, read_comb_bstd_f9, nara_acc_bstd_f9, nara_comp_bstd_f9, wold_comp_bstd_f8) %>% 
  pivot_longer(cols = c(read_comb_bstd_f9, nara_acc_bstd_f9, nara_comp_bstd_f9, wold_comp_bstd_f8),
               names_to = "task",
               values_to = "score") %>% 
  mutate(c = as.factor(c),
         task = factor(task, levels = c("read_comb_bstd_f9", "nara_acc_bstd_f9", "nara_comp_bstd_f9", "wold_comp_bstd_f8")))

# Class descriptives (Means & 1SD) 
ggplot(plot_data, aes(x = task, y = class_mean, group = c, colour = c)) +
  theme_bw() +
  geom_rect(aes(xmin = 0, xmax = Inf, ymin = -Inf, ymax = 85), fill = "grey95", colour = "white") +
  geom_point(aes(shape = c), position = position_dodge(0.2)) + 
  geom_line(aes(linetype = c), position = position_dodge(0.2)) + 
  geom_errorbar(aes(ymin = class_mean - (1*sd), ymax = class_mean + (1*sd)), width = 0.1, position = position_dodge(0.2)) +
  scale_x_discrete(name = "Task", labels = c("Item\naccuracy", "Passage\naccuracy", "Reading\ncomprehension", "Listening\ncomprehension")) +
  scale_y_continuous(name = "Class mean (+/- 1SD)") +
  geom_hline(yintercept = 100, linetype = "longdash") 

#ggsave("../output/figures/SimpView_E4_1SD.tiff", dpi = 600, width = 7, height = 5)

# Raw data
ggplot(plot_data, aes(x = task, y = class_mean, group = c)) +
  theme_bw() +
  #geom_rect(aes(xmin = 0, xmax = Inf, ymin = -Inf, ymax = 85), fill = "grey95", colour = "white") +
  geom_point(data = long_data, aes(x = task, y = score, group = c, colour = c), alpha = 0.1, 
             position = position_jitterdodge(jitter.height = 1), show.legend = FALSE) + 
  geom_point(aes(shape = c), position = position_dodge(0.75)) + 
  scale_x_discrete(name = "Task", labels = c("Item\naccuracy", "Passage\naccuracy", "Reading\ncomprehension", "Listening\ncomprehension")) +
  scale_y_continuous(name = "Individual scores") +
  geom_hline(yintercept = 100, linetype = "longdash")

#ggsave("../output/figures/SimpView_E4_raw.tiff", dpi = 600, width = 7, height = 5)

# Scatterplot - item accuracy by reading comprehension
ggplot(std_data, aes(read_comb_bstd_f9, nara_comp_bstd_f9, color = as.factor(c))) + 
  theme_bw() +
  geom_point() + 
  geom_hline(yintercept = 85, linetype = "longdash") +
  geom_vline(xintercept = 85, linetype = "longdash") + 
  xlab("Item accuracy") + 
  ylab("Reading comprehension")
#ggsave("../output/figures/SimpView_E4_itemXcomp.tiff", dpi = 600, width = 7, height = 5)

# Scatterplot - passage accuracy by passage comprehension
ggplot(std_data, aes(nara_acc_bstd_f9, nara_comp_bstd_f9, color = as.factor(c))) + 
  theme_bw() +
  geom_point() + 
  geom_hline(yintercept = 85, linetype = "longdash") +
  geom_vline(xintercept = 85, linetype = "longdash") + 
  xlab("Passage accuracy") +
  ylab("Reading comprehension")
#ggsave("../output/figures/SimpView_m2rk5_test.tiff", dpi = 600, width = 7, height = 5)
```









```{r}
# Trial - traditional criteria
all_data <- read.csv("../data/processed/WP1_data_subA.csv") %>%
  select(cidB3153, age_m_f8, age_m_f9, nara_comp_raw_f9, nara_acc_raw_f9, read_word_raw_f9, read_nonw_raw_f9, wold_comp_raw_f8,
         nara_comp_std_f9, nara_acc_std_f9) %>% 
  mutate(read_comb_raw_f9 = (read_word_raw_f9 + read_nonw_raw_f9))

std_all<- sample_std(data = all_data, test_scores = c("read_comb_raw_f9", "nara_acc_raw_f9", "nara_comp_raw_f9"), age = "age_m_f9")
std_all <- sample_std(data = std_all, test_scores = c("wold_comp_raw_f8"), age = "age_m_f8")


std_all %>% 
  filter(nara_comp_bstd_f9 < 85 & read_comb_bstd_f9 >= 85 & (read_comb_bstd_f9 - nara_comp_bstd_f9 >= 10)) %>% 
  summarise(n = n(), nara_comp = mean(nara_comp_std_f9), nara_acc = mean(nara_acc_std_f9))

poorcomp <- std_all %>% 
  filter(nara_comp_bstd_f9 < 85 & read_comb_bstd_f9 >= 85 & (read_comb_bstd_f9 - nara_comp_bstd_f9 >= 10)) %>% 
  select(cidB3153) %>% 
  mutate(pc_threshold = 1) 

poorcomp_long <- long_data %>% 
  right_join(poorcomp, by = "cidB3153") %>% 
  mutate(c = as.factor(c),
         task = factor(task, levels = c("read_comb_bstd_f9", "nara_acc_bstd_f9", "nara_comp_bstd_f9", "wold_comp_bstd_f8")))

poorcomp_long %>% 
  distinct(cidB3153, .keep_all = TRUE) %>% 
  group_by(c) %>% 
  count()
  
# Raw data
ggplot(plot_data, aes(x = task, y = class_mean, group = c)) +
  theme_bw() +
  #geom_rect(aes(xmin = 0, xmax = Inf, ymin = -Inf, ymax = 85), fill = "grey95", colour = "white") +
  geom_point(aes(shape = c), position = position_dodge(0.75)) + 
  geom_point(data = filter(long_data, !(cidB3153 %in% poorcomp_long$cidB3153)), 
             aes(x = task, y = score, fill = c, group = c, colour = c), alpha = 0.01, 
             position = position_jitterdodge(jitter.height = 1), show.legend = FALSE) + 
  #ggnewscale::new_scale_colour() + 
  geom_point(data = poorcomp_long, aes(x = task, y = score, group = c, fill = c, colour = c), alpha = 0.5, 
             position = position_jitterdodge(jitter.height = 1), show.legend = FALSE) + 
  #scale_colour_manual(values = rep("black", 4)) +
  scale_x_discrete(name = "Task", labels = c("Item\naccuracy", "Passage\naccuracy", "Reading\ncomprehension", "Listening\ncomprehension")) +
  scale_y_continuous(name = "Individual scores") +
  geom_hline(yintercept = 100, linetype = "longdash") 


#ggsave("../output/figures/SimpView_E4_raw_PC.tiff", dpi = 600, width = 7, height = 5)



```
