---
title: 'WP1: Simple View - Final Class Model'
output: 
  html_document:
    toc: true
    toc_float: true
    
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output/analysis/") })
---

This script is used for extracting descriptive statistics for the final class model (LPA or FMM).

# Set-up

```{r libraries}
# Specify packages
pkgs <- c("tidyverse", "MplusAutomation", "texreg", "kableExtra")

# Use groundhog package to load correct package versions
# if (!require("groundhog")) install.packages("groundhog")
# groundhog.day <- "2021-03-01"
# groundhog::groundhog.library(pkgs, groundhog.day)

# Or if not using groundhog package - load currently installed package versions separately (reproducibility not guaranteed)
invisible(lapply(pkgs, library, character.only = TRUE))
```

Read in selected model file that contains class assignment data, and data file of descriptive statistics. Merge and check. 

```{r read-data}
# Read in and extract class data
final_mod <- readModels(target = "./mplus_models/fmm", filefilter = "a_rerun_3class")  # final model file
class_data <- as.data.frame(final_mod$savedata) %>% 
  rename_all(tolower) %>% 
  select(yp_no, acc:c, f8age, f9age)  # keep age data for additional checking during merge

# Read in and extract relevant variables from data file
# ppt_data <- read.csv("../data/processed/WP1_data.csv") %>% 
#   select(yp_id, age_m_f8, age_m_f9, nara_comp_raw_f9, nara_acc_raw_f9, read_word_raw_f9, read_nonw_raw_f9, wold_comp_raw_f8,
#          nara_comp_std_f9, nara_acc_std_f9)

### ... simulated version for testing
ppt_data <- read.csv("../data/simulated/WP1_SimpView_sim_raw_n1000_k3.csv") %>%
  select(yp_id, yp_no, age_m_f8, age_m_f9, nara_comp_raw_f9, nara_acc_raw_f9, read_word_raw_f9, read_nonw_raw_f9, wold_comp_raw_f8) %>% 
  mutate(nara_comp_std_f9 = round(rnorm(n(), 100, 15)), # Create dummy variables for now
         nara_acc_std_f9 = round(rnorm(n(), 100, 15)))  

# Merge
ppt_classes <- ppt_data %>% 
  full_join(class_data, by = c("yp_no",
                             "age_m_f8" = "f8age",
                             "age_m_f9" = "f9age"))
```


# Whole sample descriptives

```{r all-desc}
ppt_classes %>% 
  select(-yp_id, -yp_no, -c(acc:c)) %>% 
  describe() %>% 
  as.data.frame() %>% 
  select(n, mean, sd, min, max) %>% 
  round(., 2)
```

The NARA comes with test-normed standardised scores, but the others have raw score only (bespoke tests and/or only subset of items). Compute sample standardised scores, accounting for age variability.

```{r standardise}
sample_std <- function(test_scores = NA, age = NA, data = NA){
  
  # Create row references
  data <- data %>% 
    rownames_to_column("rowno_id")
  
  # For each variable in list
  for (test in test_scores){
    
      # Create new variable name
      var_name = str_replace(test, "raw", "bstd")   # for bespoke standard
    
      # Regress out age
      lm_age <- lm(get(test) ~ get(age), data = data)
      
      # Use standardised residuals, and then convert to standardised score (M = 100, SD = 15)
      ppt_resid <- rstandard(lm_age)
      std_score <- round(ppt_resid*15 + 100)

      # Convert new standardised score to data frame
      append_score <- data.frame(std_score) %>%
        set_names(var_name) %>%
        rownames_to_column("rowno_id")

      # Merge score by row reference
      data <- data %>%
        left_join(append_score, by = "rowno_id") 
  }
  
  # Remove row reference from final output
  data <- data %>% 
    select(-rowno_id)
}

std_data <- sample_std(data = ppt_classes, test_scores = c("read_nonw_raw_f9", "read_word_raw_f9", "nara_acc_raw_f9", "nara_comp_raw_f9"), age = "age_m_f9")
std_data <- sample_std(data = std_data, test_scores = c("wold_comp_raw_f8"), age = "age_m_f8")
```


Intend to report test-standardised scores for the NARA, but check how the two compare. 
```{r std-compare}
# Accuracy
plot(std_data$nara_acc_std_f9, std_data$nara_acc_bstd_f9) +
  abline(1, 1)

# Comprehension
plot(std_data$nara_comp_std_f9, std_data$nara_comp_bstd_f9) +
  abline(1, 1)
```

# Class descriptives 

```{r class-desc}
class_averages_raw <- ppt_classes %>% 
  group_by(c) %>% 
  summarise_at(vars(read_nonw_raw_f9, read_word_raw_f9, nara_acc_raw_f9, nara_comp_raw_f9, wold_comp_raw_f8 ), 
               list(mean = mean, sd = sd, min = min, max = max), na.rm = TRUE) %>% 
  round(.,2)

class_averages_raw

class_averages_std <- std_data %>% 
  group_by(c) %>% 
  summarise_at(vars(read_nonw_bstd_f9, read_word_bstd_f9, nara_acc_std_f9, nara_comp_std_f9, wold_comp_bstd_f8), 
               list(mean = mean, sd = sd, min = min, max = max), na.rm = TRUE) %>% 
  round(.,2)

class_averages_std
```

```{r class-plot}
#### STILL TO-DO / TO MAKE NICE ####
# would be nice to add distributions, but might end up looking messy

# Summarise for plotting
plot_data <- std_data %>% 
  select(yp_id, c, read_nonw_bstd_f9, read_word_bstd_f9, nara_acc_std_f9, nara_comp_std_f9, wold_comp_bstd_f8) %>% 
  pivot_longer(cols = c(read_nonw_bstd_f9, read_word_bstd_f9, nara_acc_std_f9, nara_comp_std_f9, wold_comp_bstd_f8),
               names_to = "task",
               values_to = "score") %>% 
  group_by(c, task) %>% 
  summarise(class_mean = mean(score, na.rm = TRUE), sd = sd(score, na.rm = TRUE), se = sd/sqrt(n())) %>% 
  mutate(c = as.factor(c),
         task = factor(task, levels = c("read_nonw_bstd_f9", "read_word_bstd_f9", "nara_acc_std_f9", "nara_comp_std_f9", "wold_comp_bstd_f8")))

ggplot(plot_data, aes(x = task, y = class_mean, group = c, colour = c)) +
  theme_bw() +
  geom_point(aes(shape = c)) + 
  geom_line(aes(linetype = c)) + 
  geom_errorbar(aes(ymin = class_mean - (2*se), ymax = class_mean + (2*se)), width = 0.1)
  scale_x_discrete()

```
