---
title: 'WP1: Data extraction'
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r libraries}
# Note - no up-to-date version of packages with MCAR tests, install prev versions
#devtools::install_version("MissMech", version="1.0.2", repos=getOption("repos")["CRAN"])
#devtools::install_version("BaylorEdPsych", version="0.5", repos=getOption("repos")["CRAN"])
#devtools::install_version("mvnmle", version="0.1-11.1", repos=getOption("repos")["CRAN"])

library(tidyverse)
library(readxl)
library(naniar)
library(finalfit)
library(MissMech)
library(BaylorEdPsych)
library(broom)

options(scipen = 999)
```

This script incorporates the selection of relevant variables and cases for the research questions included in Work Package 1: *Identifying heterogeneity in reading comprehension difficulties using data collected from ALSPAC*. It also includes preliminary exploration and documentation of patterns of missingness in the key variables of interest.

# Access statement

Under ALSPAC's Data Access arrangements, we will receive all data associated with the grant at the beginning of the project. However, the grant incorporates a number of sub-projects (or work packages) that build on each other, making it challenging to specify later all hypotheses and analysis plans in advance.

We will pre-register each set of hypotheses and analyses prior to working with the data for each sub-project, and make clear in the pre-registration where analyses include variables that we have already interacted with. We will extract the relevant data for each sub-project only after it has been pre-registered, and work only with the relevant data for that sub-project.

# Extracting data

### Rename variables 

Read in and rename variables (might need some tweaking depending on format data arrives in)

```{r rename-data}
# Read in data
data_orig <- read.csv("../data/simulated/dummy_alspac.csv")
#data_orig <- read.csv("../data/raw/XXXXXX.csv")

# Read in codebook 
codebook <- read_excel("../data/codebook.xlsx", sheet = "codebook") %>% 
  select(`Var name`, R_var_name)

# Rename
data_named <- data_orig 
names(data_named) <- codebook$R_var_name[match(names(data_named), codebook$`Var name`)]
```

### WP1 Dataset 

Rename variables received from ALSPAC according to the codebook; select relevant variables; select participants to be included (i.e., those who completed the Neale Analysis of Reading Ability at age 9); create sum scores to be used.

```{r wp1-vars}
# Variables for pre-registered analyses
RQ1 <- c("age_m_f8", "age_m_f9",                                      # age
         "read_word_raw_f9", "read_nonw_raw_f9", "nara_acc_raw_f9",    # accuracy
         "nara_comp_raw_f9", "wold_comp_raw_f8")                       # comprehension

RQ2 <- c("age_m_f10",                                                # age
         "nara_rate_raw_f9",                                         # fluency
         "wold_vcb_strt_f8", "wold_vcb_i1_f8", "wold_vcb_i2_f8", "wold_vcb_i3_f8", "wold_vcb_i4_f8", "wold_vcb_i5_f8", 
         "wold_vcb_i6_f8", "wold_vcb_i7_f8", "wold_vcb_i8_f8", "wold_vcb_i9_f8", "wold_vcb_i10_f8",  "wisc_vcb_raw_f8",     # vocabulary
         "wisc_pcmp_raw_f8", "wisc_code_raw_f8", "wisc_parr_raw_f8", "wisc_bloc_raw_f8", "wisc_obja_raw_f8",  # nonverbal ability 
         "wisc_bwsp_raw_f8", "cntsp_span_raw_f10",                     # working memory
         "teach_slct_raw_f8", "teach_divd_raw_f8", "teach_ctrs_raw_f8", "teach_ctro_raw_f8", "sdq_hyp_prnt_ku", "sst_250a_raw_f10", "sst_150a_raw_f10") #attention  

# Variables for descriptives, exploratory analyses (subject to change)
other <- c("nara_acc_std_f9", "nara_comp_std_f9", "nara_rate_std_f9", "nara_npass_f9",
           "wisc_vcb_std_f8",
           "wisc_pcmp_std_f8", "wisc_code_std_f8", "wisc_parr_std_f8", "wisc_bloc_std_f8", "wisc_obja_std_f8", "wisc_perf_std_f8") 
  
# Additional variables for exploring missingness
missingness <-  c("sex", "m_cigs_n_a", "m_marry_a", "m_socc_c", "m_fdif_tot_c", "m_home_a")

# All variables for extraction
wp1_vars <- c("yp_id", RQ1, RQ2, other, missingness)  
```

Extract and process
```{r wp1-data}
wp1_data <- data_named %>% 
  
  # Select relevant variables
  select(all_of(wp1_vars)) %>% 
  
  # Filter for participants with NARA
  filter(!is.na(nara_comp_raw_f9)) %>% 
  
  # Create processed vars 
  mutate(wold_vcb_raw_f8 = ifelse(wold_vcb_strt_f8 == 1,                      # wold vocab sum score
                                  rowSums(select(., contains("wold_vcb_i"))), # TO-DO check missing item-level data
                                  NA),
         teach_ctr_diff_f8 = teach_ctro_raw_f8 - teach_ctrs_raw_f8) %>%      # difference score for teach attentional control

  # Remove unnecessary variables
  select(-c(wold_vcb_strt_f8:wold_vcb_i10_f8), -c(teach_ctro_raw_f8:teach_ctrs_raw_f8))
 
# Test whether meets expected  number of participants
length(unique(wp1_data$yp_id)) == 6943

# Save data
# write.csv(wp1_data, "../data/processed/WP1_data.csv", row.names = FALSE)
```

**TO-DO** Process missing data labels? 

# Missingness

### Description

Document prevalence and patterns of missingness for variables to be included in the pre-registered analyses.

```{r missing-descr}
# Prevalence
wp1_data %>% 
  select(-all_of(c(other, missingness))) %>% 
  vis_miss()

# Pattern
wp1_data %>% 
  select(-all_of(c(other, missingness))) %>% 
  missing_pattern()
```

### Structure

Use two types of test to examine whether data is likely MCAR. Little's (1988) MCAR test examines whether variable means differ between missing data patterns. Jamshidian & Jalal's (2010) test examines whether the variances and covariances differ betwen missing data patterns. 

```{r missing-tests}
# Little's MCAR test 
missingness_little <- wp1_data %>% 
  select(-all_of(c(other))) %>% 
  select(where(is.numeric)) %>% 
  LittleMCAR()

missingness_little$chi.square
missingness_little$df
missingness_little$p.value

# Jamshidian & Jalal (2010) MCAR test
wp1_data %>% 
  select(-all_of(c(other))) %>% 
  select(where(is.numeric)) %>% 
  TestMCARNormality()
```

### Predictors of missingness 

Given the assessments are all taken from assessments close in time, we do not anticipate structured missingness in these data. 

```{r missing-predict}
# Create updated list of pre-registered variables (incorporating new vars created above)
prereg_vars <- names(wp1_data) %>% 
  setdiff(c(missingness, other))

# Create data frame for storing significant predictors
miss_pred <- data.frame(missing_val = as.character(), predictor = as.character(), estimate = as.numeric(), std.error = as.numeric(), statistic = as.numeric(), p.value = as.numeric(), p.holm = as.numeric())

for (var in prereg_vars){
  
# for (var in names(sim_data)) { (used for testing)
  
  # Create data frame with selected variable marked for missingness, remove irrelevant columns 
  data <- wp1_data %>% 
    select(-all_of(c(other))) %>% 
    mutate(is_missing = ifelse(is.na(get(var)) == TRUE, 1, 0)) %>% 
    select(-c(yp_id, var))
  
  # If no missing data, print statement
  if (sum(data$is_missing) == 0) {
      print(paste0("No missing values for variable ", var))
    
  # Else if missing data present...
  } else {
      
      # Test every other variable as a predictor
      missing_model <- glm(is_missing ~ ., family = binomial("logit"), data = data)
      tidy_model <- tidy(missing_model) %>%  # format output
        mutate(missing_val = var, p.adjust = p.adjust(p.value, "holm")) %>% 
        rename(predictor = term) %>% 
        select(missing_val, predictor, estimate, std.error, statistic, p.value, p.adjust)
      
      # If any of the values are significant, print output labelled with predictor and variable
      if(any(tidy_model$p.value < .05)) {
        print(paste0("Potential predictors (non-corrected) of missing values for variable  ", var))
        
        # Append significant predictors to missing data table
        sig_vals <- tidy_model %>% 
          filter(p.value < .05)
        miss_pred <- rbind(miss_pred, sig_vals)

      # Otherwise print that no significant predictors were found
      } else {
        print(paste0("No significant predictors of missing values found for variable ", var))
      }
  }
}
print(miss_pred)
```
