---
title: 'WP1: Data extraction'
output: 
  html_document:
    toc: true
    toc_float: true
date: '24/05/2021'

knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output/analysis/") })
---

This script incorporates the selection of relevant variables and cases for the research questions included in Work Package 1: *Identifying heterogeneity in reading comprehension difficulties using data collected from ALSPAC*. It also includes preliminary exploration and documentation of patterns of missingness in the key variables of interest, and splits the sample in two for calibration and validation purposes.

```{r libraries, message = FALSE, warning = FALSE}
# Note - no up-to-date version of packages with MCAR tests, install prev versions - use with caution
#devtools::install_version("MissMech", version="1.0.2", repos=getOption("repos")["CRAN"])
#devtools::install_version("BaylorEdPsych", version="0.5", repos=getOption("repos")["CRAN"])
#devtools::install_version("mvnmle", version="0.1-11.1", repos=getOption("repos")["CRAN"])

library(data.table)
library(tidyverse)
library(readxl)
library(naniar)
library(finalfit)
library(MissMech)
library(BaylorEdPsych)
library(broom)
library(psych)
library(WriteXLS)
library(knitr)

options(scipen = 999)
```

```{r create-dir}
# Create subdirectories for storing data files and output, if do not already exist
if(dir.exists("./output/")==FALSE){dir.create("./output/")}
if(dir.exists("./output/figures/")==FALSE){dir.create("./output/figures/")}
if(dir.exists("./output/analysis/")==FALSE){dir.create("./output/analysis/")}
if(dir.exists("./data/processed/")==FALSE){dir.create("./data/processed/")}
```


# Access statement

Under ALSPAC's Data Access arrangements, we will receive all data associated with the grant at the beginning of the project. However, the grant incorporates a number of sub-projects (or work packages) that build on each other, making it challenging to specify all later hypotheses and analysis plans in advance.

We will pre-register each set of hypotheses and analyses prior to working with the data for each sub-project, and make clear in the pre-registration where analyses include variables that we have already interacted with. We will extract the relevant data for each sub-project only after it has been pre-registered, and work only with the relevant data for that sub-project.

The pre-registration for this sub-project can be found at [https://osf.io/4zahf](https://osf.io/4zahf).

# Extracting data

### Rename variables 

Read in and rename variables

```{r rename-data}
# Read in data
#data_orig <- read.csv("../data/simulated/dummy_alspac.csv")  # for script testing
data_orig <- fread("../data/raw/B3153/Nation_19May21.csv") # ALSPAC data

# Read in codebook for renaming
codebook <- read_excel("../data/codebook.xlsx", sheet = "codebook") %>% 
  filter(workbook != "education" | is.na(workbook)) %>%   # education data coming from separate database
  select(`Var name`, R_var_name)

# # List variables not in existing codebook
# setdiff(codebook$`Var name`, names(data_orig))
# setdiff(names(data_orig), codebook$`Var name`)
# 
# # Extract variable info for updating codebook
# alspac_codebook <- read_excel("../data/raw/B3153/Variable information.xlsx", skip = 1) %>% 
#   filter(!Variable %in% codebook$`Var name`)
# write.csv(alspac_codebook, "../data/raw/codebook_update.csv")

# Rename
data_named <- data_orig 
names(data_named) <- codebook$R_var_name[match(names(data_named), codebook$`Var name`)]

# Create unique identifier per participant
data_named <- data_named %>% 
  mutate(yp_id = paste0(cidB3153, qlet))
```

### Participant numbers

```{r dataset-nos}
# Number of participants - minus triplets/quadruplets (n = 13 individuals)
nrow(data_named)

# Number of pregnancies
length(unique(data_named$cidB3153))

# Participants alive at 1 year (ALSPAC baseline group)
data_named %>% 
  mutate(alive_1yr = ifelse(kz011b == -9999, "withdrawn",
                            ifelse(kz011b == 2, 0, kz011b))) %>% 
  group_by(alive_1yr) %>% 
  count()

# Participants in core sample
data_named %>% 
  mutate(in_core = ifelse(in_core == -9999, "withdrawn",
                            ifelse(in_core == 2, 0,
                                   ifelse(in_core == -1, "birth outcome not known", in_core)))) %>% 
  group_by(in_core) %>% 
  count()

# Core sample alive at 1 year (ALSPAC baseline group)
data_named %>% 
  filter(in_core == 1) %>% 
  mutate(alive_1yr = ifelse(kz011b == -9999, "withdrawn",
                            ifelse(kz011b == 2, 0, kz011b))) %>% 
  group_by(alive_1yr) %>% 
  count()

# Participants enrolled as part of phase 2, during focus@7
data_named %>% 
  mutate(in_phase2 = ifelse(in_phase2 == -9999, "withdrawn",
                            ifelse(in_phase2 == 2, 0,
                                   ifelse(in_phase2 == -1, "birth outcome not known", in_phase2)))) %>% 
  group_by(in_phase2) %>% 
  count()

# Participants enrolled as part of phase 3, between focus@7 and age 18
data_named %>% 
  mutate(in_phase3 = ifelse(in_phase3 == -9999, "withdrawn",
                            ifelse(in_phase3 == 2, 0,
                                   ifelse(in_phase3 == -1, "birth outcome not known", in_phase3)))) %>% 
  group_by(in_phase3) %>% 
  count()

# Participants enrolled as part of phase 4, between ages 18 and 24
data_named %>% 
  mutate(in_phase4 = ifelse(in_phase4 == 2, 0, in_phase4)) %>% 
  group_by(in_phase4) %>% 
  count()
```

### WP1 Variables 

Specify variables to be used in these analyses.

```{r wp1-vars}
# Variables for pre-registered analyses
RQ1 <- c("age_m_f8", "age_m_f9",                                       # age
         "read_word_raw_f9", "read_nonw_raw_f9", "nara_acc_raw_f9",    # accuracy
         "nara_comp_raw_f9", "wold_comp_raw_f8")                       # comprehension

RQ2 <- c("age_m_f10",                                                # age
         "nara_rate_raw_f9",                                         # fluency
         "wold_vcb_strt_f8", "wold_vcb_i1_f8", "wold_vcb_i2_f8", "wold_vcb_i3_f8", "wold_vcb_i4_f8", "wold_vcb_i5_f8", 
         "wold_vcb_i6_f8", "wold_vcb_i7_f8", "wold_vcb_i8_f8", "wold_vcb_i9_f8", "wold_vcb_i10_f8",  "wisc_vcb_raw_f8",     # vocabulary
         "wisc_pcmp_raw_f8", "wisc_code_raw_f8", "wisc_parr_raw_f8", "wisc_bloc_raw_f8", "wisc_obja_raw_f8",  # nonverbal ability 
         "wisc_bwsp_raw_f8", "cntsp_span_raw_f10",                     # working memory
         "teach_slct_raw_f8", "teach_divd_raw_f8", "teach_ctrs_raw_f8", "teach_ctro_raw_f8", "sdq_hyp_prnt_ku") #attention  

# Variables for descriptives, exploratory analyses (subject to change)
other <- c("yp_id", "cidB3153", "qlet",
           "sex",
           "nara_acc_std_f9", "nara_comp_std_f9", "nara_rate_std_f9", "nara_npass_f9",
           "wisc_vcb_std_f8",
           "wisc_pcmp_std_f8", "wisc_code_std_f8", "wisc_parr_std_f8", "wisc_bloc_std_f8", "wisc_obja_std_f8", "wisc_perf_std_f8") 
  
# Additional variables for exploring missingness
missingness <-  c("ethn_c", "wght_birth", "gest_birth", "m_age_birth", "m_edu_c", "m_home_a", "mtcq_vcb_raw_kg")

# All variables for extraction
wp1_vars <- c(RQ1, RQ2, other, missingness)  
```

Extract relevant variables (including all observations at this stage) and recode missing data values.

```{r wp1-all-obs}
# Extract relevant variables from dataset, and processing missing data
wp1_allobs <- data_named %>% 
  
  # Select relevant variables
  select(all_of(wp1_vars)) %>% 
  
  # Relabel missing data 
  replace_with_na_at(.vars = wp1_vars[!wp1_vars %in% c("teach_divd_raw_f8", "teach_slct_raw_f8")],  # exclude TEACh variables as have true negatives
                       condition = ~.x < 0) %>% 
  
  replace_with_na(replace = list(teach_divd_raw_f8 = c(-9999, -103, -101, -2),
                                 teach_slct_raw_f8 = c(-9999, -103, -102, -101)))


# Inspect summary statistics and check that match up with expectations from data dictionary (i.e., to make sure missing data recoded correctly)
describe(wp1_allobs) %>%
  round(., 2) %>% 
  kable() 
```

```{r var-pre-checks, echo = FALSE, eval = FALSE}
# PRE-CHECKS FOR ITEMS USED TO DERIVE NEW VARIABLES

# WOLD VOCABULARY

### Check missing item level data
wp1_allobs %>% 
  filter(wold_vcb_strt_f8 == 1) %>% 
  select(contains("wold_vcb_i")) %>% 
  vis_miss()

### Recode incorrect items to 0
test <- wp1_allobs %>% 
  filter(wold_vcb_strt_f8 == 1) %>% 
  select(contains("wold_vcb_i")) %>% 
  mutate_at(vars(contains("wold_vcb_i")), ~ifelse(. == 2 | . == 9, 0, .))

# TEACH ATTENTIONAL CONTROL

### Check missing item level data
data_named %>% 
  filter(in_f08 == 1) %>% 
  select(contains("teach_ctr")) %>% 
  vis_miss()

test <- wp1_allobs %>% 
  mutate(teach_ctr_diff_f8 = teach_ctro_raw_f8 - teach_ctrs_raw_f8) %>% 
  select(contains("teach_ctr"))

hist(test$teach_ctr_diff_f8, breaks = 100)
```

### WP1 subsample

Identify twins for exclusion (not pre-registered due to an oversight).

```{r twins}
exclude <- wp1_allobs %>% 
  filter(!is.na(nara_comp_raw_f9)) %>% 
  group_by(cidB3153) %>% 
  filter(n() > 1) %>% 
  slice_sample(n = 1)

# Print number of twins for exclusion
nrow(exclude)
```

Select participants to be included (i.e., those who completed the Neale Analysis of Reading Ability at age 9, minus one of each twin pair); create sum scores to be used.

```{r wp1-data}
wp1_data <- wp1_allobs %>%  
  
  # Filter for participants with NARA
  filter(!is.na(nara_comp_raw_f9)) %>% 
  
  # Filter out twins for exclusion
  anti_join(exclude, by = c("cidB3153", "qlet")) %>% 
  
  # Pre-process WOLD item data (incorrect = 2 >> incorrect = 0)
  mutate_at(vars(contains("wold_vcb_i")), ~ifelse(. == 2 | . == 9, 0, .)) %>% 
  
  # Create processed vars 
  mutate(wold_vcb_raw_f8 = ifelse(wold_vcb_strt_f8 == 1,                      # wold vocab sum score
                                  rowSums(select(., contains("wold_vcb_i"))), 
                                  NA),
         teach_ctr_diff_f8 = teach_ctro_raw_f8 - teach_ctrs_raw_f8) %>%      # difference score for teach attentional control

  # Remove unnecessary variables
  select(-c(wold_vcb_strt_f8:wold_vcb_i10_f8), -c(teach_ctro_raw_f8:teach_ctrs_raw_f8))
 

# Inspect summary statistics
describe(wp1_data) %>%
  round(., 2) %>% 
  kable() 
```

# Missingness

### Description

Document prevalence and patterns of missingness for variables to be included in the pre-registered analyses.

```{r missing-descr}
# Prevalence
missing_prev <- wp1_data %>% 
  select(-all_of(c(other, missingness, "yp_id"))) %>%
  select(order(colnames(.))) %>% 
  vis_miss(cluster = TRUE) 

missing_prev + 
  theme(axis.text.x = element_text(angle = 50),
  plot.margin = unit(c(1, 3, 1, 1), "cm"))

ggsave("../output/figures/WP1_missingness.png")

# Pattern
# missing_pattern <- wp1_data %>% 
#   select(-all_of(c(RQ1, other, missingness, "yp_id", "age_m_f10"))) %>% 
#   missing_pattern()
#gsave("../output/figures/WP1_missingness_patterns.png", plot = missing_pattern)
```

### Structure

Use two types of test to examine whether data is likely MCAR. Little's (1988) MCAR test examines whether variable means differ between missing data patterns. Jamshidian & Jalal's (2010) test examines whether the variances and covariances differ betwen missing data patterns. 

*TO FIX*

```{r missing-tests}
# Little's MCAR test - DOES NOT RUN 
# missingness_little <- wp1_data %>% 
#   select(-all_of(c(other, "yp_id"))) %>% 
#   select(where(is.numeric)) %>% 
#   LittleMCAR()
# 
# print(cat("\nLITTLE'S TEST \nchi.square:", missingness_little$chi.square, "\ndf:", missingness_little$df, "\np = ", missingness_little$p.value))

# Jamshidian & Jalal (2010) MCAR test
wp1_data %>% 
  select(-all_of(c(other))) %>% 
  select(where(is.numeric)) %>% 
  TestMCARNormality()
```

### Predictors of missingness 

Given the assessments are all taken from assessments close in time, we do not anticipate structured missingness in these data. 

*TO FIX*

```{r missing-predict}
# Create updated list of pre-registered variables (incorporating new vars created above)
prereg_vars <- names(wp1_data) %>% 
  setdiff(c(missingness, other))

# Create data frame for storing significant predictors
miss_pred <- data.frame(missing_val = as.character(), predictor = as.character(), estimate = as.numeric(), std.error = as.numeric(), statistic = as.numeric(), p.value = as.numeric(), p.holm = as.numeric())

for (var in prereg_vars){
  
# for (var in names(sim_data)) { (used for testing)
  
  # Create data frame with selected variable marked for missingness, remove irrelevant columns 
  data <- wp1_data %>% 
    select(-all_of(c(other))) %>% 
    mutate(is_missing = ifelse(is.na(get(var)) == TRUE, 1, 0)) %>% 
    select(-c(var))
  
  # If no missing data, print statement
  if (sum(data$is_missing) == 0) {
      print(paste0("No missing values for variable ", var))
    
  # Else if missing data present...
  } else {
      
      # Test every other variable as a predictor
      missing_model <- glm(is_missing ~ ., family = binomial("logit"), data = data)
      tidy_model <- tidy(missing_model) %>%  # format output
        mutate(missing_val = var, p.adjust = p.adjust(p.value, "holm")) %>% 
        rename(predictor = term) %>% 
        select(missing_val, predictor, estimate, std.error, statistic, p.value, p.adjust)
      
      # If any of the values are significant, print output labelled with predictor and variable
      if(any(tidy_model$p.value < .05)) {
        print(paste0("Potential predictors (non-corrected) of missing values for variable  ", var))
        
        # Append significant predictors to missing data table
        sig_vals <- tidy_model %>% 
          filter(p.value < .05)
        miss_pred <- rbind(miss_pred, sig_vals)

      # Otherwise print that no significant predictors were found
      } else {
        print(paste0("No significant predictors of missing values found for variable ", var))
      }
  }
}
print(miss_pred)
```

# Split sample and save out data

Create subsamples to be used for RQ1 initial class enumeration process and subsequent validation. 

```{r split-half}
# Set seed for reproducibility
set.seed(25052021)

# Create numeric ID variable for working with mplus, remove unnecessary variables and tidy order
wp1_tidy <- wp1_data %>% 
  select(-all_of(missingness)) %>% 
  relocate(c(yp_id, cidB3153, qlet, sex), .before = 1) %>% 
  relocate(age_m_f10, .after = age_m_f9) %>% 
  relocate(teach_ctr_diff_f8, .after = teach_divd_raw_f8) %>% 
  relocate(wold_vcb_raw_f8, .after = wisc_vcb_raw_f8)

# Create exploratory/calibration dataset
sampleA <- wp1_tidy %>% 
  sample_frac(.50)

# Create confirmatory/validation dataset
sampleB <- wp1_tidy %>% 
  anti_join(sampleA, by = "yp_id")
  
# Check
sampleA$yp_id[(sampleA$yp_id %in% sampleB$yp_id)]
sampleB$yp_id[(sampleB$yp_id %in% sampleA$yp_id)]
(nrow(sampleA) + nrow(sampleB)) == nrow(wp1_tidy)

# Save
write.csv(sampleA, "../data/processed/WP1_data_subA.csv", row.names = FALSE)
write.csv(sampleB, "../data/processed/WP1_data_subB.csv", row.names = FALSE)
write.csv(wp1_tidy, "../data/processed/WP1_data_all.csv", row.names = FALSE)
```

Save record of additional variables created to the codebook. A record of generated variables will need to be returned to ALSPAC upon completion of the project.

```{r new_vars}
# List variables not in original dataset
vars_created <- setdiff(names(wp1_data), names(data_named)) %>% 
  as.data.frame() %>% 
  set_names("R_var_name") %>% 
  mutate(workpack = 1,
         script_created = "WP1_DataExtraction.Rmd")

# Read in sheet of codebook that contains created variables
new_vars <- read_excel("../data/generated_variables.xlsx", sheet = "new_vars") %>% 
  mutate(R_var_name = as.character(R_var_name))

# Combine the two lists, retain only distinct variable names 
vars_combined <- new_vars %>% 
  bind_rows(vars_created) %>% 
  distinct(R_var_name, .keep_all = TRUE) 

# Save out new sheet if new rows have been added
if (nrow(vars_combined) > nrow(new_vars)){
  
  # Update sheet
  new_vars <- vars_combined
  
  # Read in existing sheets
  metadata <- read_excel("../data/generated_variables.xlsx", sheet = "metadata")
  
  # Write out together
  WriteXLS(c("metadata", "new_vars"), "../data/generated_variables.xlsx", perl = "C:/Perl64/bin/perl.exe")
  
  # Request edits in excel
  print("New variables generated. Update documentation in generated_variables.xlsx file.")
  shell.exec(here::here("./data/generated_variables.xlsx")) # open file to add further details
}  
```

