---
title: "WP1: Profiles of comprehension difficulty - Data Pre-processing and CFA"
output: 
  html_document:
    toc: true
    toc_float: true
date: '06/05/2021'
    
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output/analysis/") })
---

This script computes basic summary statistics for the data, ensures that they align with Mplus output, and tests the factor structure to be used for subsequent analyses. This analysis is conducted using the group(s) with poor reading comprehension, extracted in WP1_SimpView_FinalModel.Rmd. 

# Set-up

```{r libraries}
# List packages
pkgs <- c("tidyverse", "psych", "naniar", "MplusAutomation", "texreg", "semPlot")

# Use groundhog package to load correct package versions
if (!require("groundhog")) install.packages("groundhog")
groundhog.day <- "2021-03-01"
groundhog::groundhog.library(pkgs, groundhog.day)

# Or if not using groundhog package - load currently installed package versions separately (faster; reproducibility not guaranteed)
# invisible(lapply(pkgs, library, character.only = TRUE))
```

```{r create-dir}
# Create subdirectories for storing mplus scripts, data files, and output, if do not already exist
if(dir.exists("./mplus_models/")==FALSE){dir.create("./mplus_models/")}
if(dir.exists("./mplus_models/compprof/")==FALSE){dir.create("./mplus_models/compprof/")}
if(dir.exists("./mplus_models/compprof/desc")==FALSE){dir.create("./mplus_models/compprof/desc")}
if(dir.exists("./mplus_models/compprof/cfa")==FALSE){dir.create("./mplus_models/compprof/cfa")} 
```

###  Load and prepare data for Mplus

Read in the original data file extracted for WP1 (contains all variables) and the classes extracted from the previous analysis. Filter dataset to include only the classes of interest for this analysis (those with poor comprehension in the context of relatively good reading accuracy). 

```{r load-data}
cp_data <- read.csv("../data/simulated/WP1_CompProf_sim_raw_n694_k3.csv")   # for script checking

# cp_data <- read.csv("../data/processed/WP1_data_all.csv")      # dataset containing all participants/variables
# groups <- read.csv("../data/processed/WP1_classes.csv") %>%    # classes extracted from SimpView model
#   select()

# cp_data <- cp_data %>%
#   left_join(groups, by = "") %>% 
#   filter(C == ) %>% 
#   select(yp_id:)
```

Mplus has an 8-character limit, so rename relevant variables with short names for modelling. 

```{r mplus-names}
# Rename any necessary, print new names
cp_data <- cp_data %>% 
  rename(naraComp = nara_comp_raw_f9,
         naraAcc = nara_acc_raw_f9,
         wordAcc = read_word_raw_f9,
         nonwAcc = read_nonw_raw_f9,
         woldComp = wold_comp_raw_f8,
         f8age = age_m_f8,
         f9age = age_m_f9,
         f10age = age_m_f10,
         naraRate = nara_rate_raw_f9,
         woldVcb = wold_vcb_raw_f8,
         wiscVcb = wisc_vcb_raw_f8,
         wiscPcmp = wisc_pcmp_raw_f8,
         wiscCode = wisc_code_raw_f8,
         wiscParr = wisc_parr_raw_f8,
         wiscBloc = wisc_bloc_raw_f8,
         wiscObja = wisc_obja_raw_f8,
         wiscBwsp = wisc_bwsp_raw_f8,
         cntspan = cntsp_span_raw_f10,
         attnSel = teach_slct_raw_f8,
         attnDiv = teach_divd_raw_f8,
         attnOpp = teach_ctr_diff_f8,
         sdqHyp = sdq_hyp_prnt_ku) 

# Check compatibility with mplus
str_sub(names(cp_data), 1, 8)
```

Variable format problematic for mplus data conversion, change ID to factor.

```{r mplus-formatting}
cp_data <- cp_data %>% 
  mutate(yp_id = as.factor(yp_id))
```


# Descriptive statistics

Extract summary statistics for each variable.

```{r summary-stats}
describe(cp_data)
vis_miss(cp_data, cluster = TRUE)
```

Patterns of missingness are more thoroughly explored for all WP1 variables in the initial data processing script. We anticipate more missingness than in the SimpView analysis as there were more measures collected from across three different clinics. 

### Check that mplus reading data as expected

Extract descriptive statistics as check.

```{r mplus-desc}
# Specify model
m_desc <- mplusObject(
  TITLE = "Data check - Descriptive statistics;",
  ANALYSIS = "type = basic;",
  usevariables = colnames(cp_data[,!names(cp_data) %in% c("yp_id", "yp_no", "k")]),
  rdata = cp_data)

# Fit model 
m_desc_fit <- mplusModeler(m_desc,
                            dataout = "./mplus_models/compprof/desc/cp_sim.dat",
                            modelout = "./mplus_models/compprof/desc/cp_check.inp",
                            check = TRUE, run = TRUE, hashfilename = TRUE)

# Read mplus output
m_desc_out <- readModels("./mplus_models/compprof/desc/cp_check.out")

# Check that sample sizes and means match descriptive statistics from above
r_summ <- describe(cp_data) %>%
  as.data.frame() %>% 
  slice(3:24) %>% 
  select(n, mean) %>% 
  round(., 2)
mplus_summ <- m_desc_out$sampstat$univariate.sample.statistics %>% 
  as.data.frame() %>%
  round(.,2) %>% 
  select(`Sample Size`, Mean) %>% 
  set_names(c("n", "mean"))

if (any((r_summ - mplus_summ) > abs(0.01))){
  print("WARNING: DIFFERENCES DETECTED BETWEEN R AND MPLUS DATA SUMMARIES. Inspect output.")
  setdiff(mplus_summ, r_summ)
  } else{
      print("Data check passed")
  }
```

# Confirmatory Factor Analysis

### Testing the factor structure

Fit the proposed factorial model.

```{r cfa-initial-fit}
# Proposed factor structure: 5 factors + rate
m_cfa1 <- mplusObject(
  TITLE = "Confirmatory Factor Analysis - 5-factor + rate;",
  ANALYSIS = "estimator = mlr; type = general;",
  MODEL = "wordAcc nonwAcc naraAcc naraComp naraRate on f9age;
            woldComp woldVcb wiscVcb wiscPcmp wiscCode wiscParr 
            wiscBloc wiscObja wiscBwsp attnSel attnDiv attnOpp on f8age;
            cntSpan on f10age;
            acc by nonwAcc wordAcc naraAcc; 
            comp by naraComp woldComp;
            naraRate;
            vocab by wiscVcb woldVcb ;
            perfiq by wiscPcmp wiscCode wiscParr wiscBloc wiscObja;
            execfun by wiscBwsp cntSpan attnSel attnDiv attnOpp sdqHyp;
            acc with comp; acc with naraRate; acc with vocab; 
            acc with perfiq; acc with execfun;
            comp with naraRate; comp with vocab; 
            comp with perfiq; comp with execfun;
            naraRate with vocab; naraRate with perfiq; 
            naraRate with execfun; 
            vocab with perfiq; vocab with execfun;
            perfiq with execfun;",
  OUTPUT = "TECH1; TECH4; stdyx; modindices;",
  PLOT = "TYPE = PLOT3;",
  usevariables = colnames(cp_data[,!names(cp_data) %in% c("yp_id", "yp_no")]),
  rdata = cp_data)

m_cfa1_fit <- mplusModeler(m_cfa1,
                           modelout = "./mplus_models/compprof/cfa/cp_cfa1.inp",
                           check = TRUE, run = TRUE)


# Inspect initial model fit
cfa1_out <- readModels(target = "./mplus_models/compprof/cfa", filefilter = "cp_cfa")
SummaryTable(cfa1_out, keepCols = c("Title", "Parameters", "LL", "CFI", "TLI", "AIC", "BIC", "RMSEA_Estimate", "RMSEA_pLT05", "SRMR"))
```


### Improving fit 

```{r cfa-inspection}
# Inspect model (by = loadings, with = covariances)
cfa1_out$warnings
cfa1_out$parameters$unstandardized
cfa1_out$parameters$stdyx.standardized

# Inspect modification indices
cfa1_out$mod_indices %>% 
  filter(operator == "WITH") %>% 
  arrange(desc(MI))
```

* Check for very large standardised residuals.  
* No accepted view on how large MIs should be. Change only one at a time, starting with largest, but must be theoretically meaningful.  
* As well as the additional modifications from the SimpView analysis, we anticipate potential covariances between aspects of executive function that are more closely related to each other, such as the working memory measures or subtests from the TEACh battery. 

E.g...

```{r cfa-modification1}
# TWO-FACTOR MODEL with additional covariance between the two working memory measures
# m_cfa2 <- update(m_cfa1,
#                  TITLE = ~ "Confirmatory Factor Analysis - Two-Factor - modification1 wm;",
#                  MODEL = ~. + "cntSpan with wiscBwsp;")
# 
# m_cfa2_fit <- mplusModeler(m_cfa2,
#                            modelout = "./mplus_models/compprof/cfa/cp_cfa2.inp",
#                            check = TRUE, run = TRUE)
# 
# # COMPARE MODELS
# cfa_models <- readModels(target = "./mplus_models/compprof/cfa", filefilter = "cp_cfa")
# SummaryTable(cfa_models, keepCols = c("Title", "Parameters", "LL", "CFI", "TLI", "AIC", "BIC", "RMSEA_Estimate", "RMSEA_pLT05", "SRMR"))
# 
# # Inspect model (by = loadings, with = covariances)
# cfa_models$cp_cfa2.out$warnings
# cfa_models$cp_cfa2.out$parameters$unstandardized
# cfa_models$cp_cfa2.out$parameters$stdyx.standardized
# 
# # Inspect modification indices
# cfa_models$cp_cfa2.out$mod_indices %>% 
#   filter(operator == "WITH") %>% 
#   arrange(desc(MI))
```


The following rules of thumb are often used to indicate a good-fitting model:
* CFI & TLI > .95  
* RMSEA <= .05  
* SRMR < .05


### CFA figure

```{r cfa-diagram}
#semPaths(semPlotModel(cfa_models$cp_cfa2.out, mplusStd = "stdyx"), what = "paths", whatLabels = "std", rotation = 2, intercepts = FALSE)
```

