---
title: "WP1: Simple View - Data Simulation"
output: html_document
---

This script simulates data designed around the variables available in ALSPAC for the first analysis (extracting sub-populations of readers in line with the Simple View of reading).

```{r libraries}
# List packages
pkgs <- c("SIN", "MASS", "psych", "tidyverse", "naniar", "GGally")

# Use groundhog package to load correct package versions
# if (!require("groundhog")) install.packages("groundhog")
# groundhog.day <- "2021-03-01"
# groundhog::groundhog.library(pkgs, groundhog.day)

# Or if not using groundhog package - load currently installed package versions separately (reproducibility not guaranteed)
invisible(lapply(pkgs, library, character.only = TRUE))

# Set random seed
set.seed(4926)
```


### Specify variables and relationships

Create dataset similar to one that we will use - three reading accuracy measures, a reading comprehension measure and listening comprehension.  

* NARA comp - range 0-45, M = 25
* NARA acc (errors) - range 0-100, M = 66  (note that a lower score is better)
* Word and nonword acc - range 0-10 
* WOLD comp - range 2-14

```{r measures}

################# CLASS 1 MEASURES ###################

# NARA - comprehension
nara_comp_m1 <- 35
nara_comp_sd1 <- 10
  
# NARA - accuracy
nara_acc_m1 <- 30
nara_acc_sd1 <- 6

# Word/nonword reading
word_acc_m1 <- 9
word_acc_sd1 <- 1
nonword_acc_m1 <- 8
nonword_acc_sd1 <- 1

# WOLD listening
wold_comp_m1 <- 10
wold_comp_sd1 <- 3


################# CLASS 2 MEASURES ###################

# NARA - comprehension
nara_comp_m2 <- 10
nara_comp_sd2 <- 5
  
# NARA - accuracy
nara_acc_m2 <- 60
nara_acc_sd2 <- 10

# Word/nonword reading
word_acc_m2 <- 4
word_acc_sd2 <- 2
nonword_acc_m2 <- 3
nonword_acc_sd2 <- 2

# WOLD listening
wold_comp_m2 <- 5
wold_comp_sd2 <- 3



################# CLASS 3 MEASURES ###################

# NARA - comprehension
nara_comp_m3 <- 20
nara_comp_sd3 <- 6
  
# NARA - accuracy
nara_acc_m3 <- 50
nara_acc_sd3 <- 10

# Word/nonword reading
word_acc_m3 <- 7
word_acc_sd3 <- 2
nonword_acc_m3 <- 5
nonword_acc_sd3 <- 2

# WOLD listening
wold_comp_m3 <- 7
wold_comp_sd3 <- 3


################# CORRELATIONS ###################

# Assumption is that we have same correlation among measures for each class (note NARA acc measures neg correlated; lower score = better)

nara_comp_acc <- -0.5
nara_comp_word <- 0.4
nara_comp_nonword <- 0.3
nara_comp_wold <- 0.6
nara_acc_word <- -0.6
nara_acc_nonword <- -0.5
nara_acc_wold <- -0.2
word_acc_nonword <- 0.7
word_acc_wold <- 0.2
nonword_acc_wold <- 0.2
```

Dataset will need to take age into account. Do not expect large age differences between groups, so specify same means and correlations for each. 
*Note - originally used same means/SDs for all 3 groups, tried separating in case source of the problem* 

```{r ages}
# Set age means - roughly guided by those that attended clinic
age_months_f8_m <- 103
age_months_f8_sd <- 4
age_months_f9_m <- 119
age_months_f9_sd <- 4

# age_months_f8_m1 <- 104
# age_months_f8_sd1 <- 4
# age_months_f9_m1 <- 121
# age_months_f9_sd1 <- 4
# 
# age_months_f8_m2 <- 101
# age_months_f8_sd2 <- 3
# age_months_f9_m2 <- 117
# age_months_f9_sd2 <- 4
# 
# age_months_f8_m3 <- 103
# age_months_f8_sd3 <- 4
# age_months_f9_m3 <- 119
# age_months_f9_sd3 <- 4


# F8 age should only correlate with measures taken at F8 
nara_comp_f8age <- 0
nara_acc_f8age <- 0
word_acc_f8age <- 0
nonword_acc_f8age <- 0
wold_f8age <- 0.1

# F9 age should only correlate with measures taken at F9
nara_comp_f9age <- 0.1
nara_acc_f9age <- -0.1
word_acc_f9age <- 0.1
nonword_acc_f9age <- 0.1
wold_f9age <- 0

# Age measures should not correlate with each other
f8age_f9age <- 0
```

### Simulate dataset

Compute covariance matrix based on the above estimates, and use to simulate specified number of observations.

```{r simulate-data}
################# COVARIANCES ###################

# List names in order
test_names <- c("nara_comp", "nara_acc", "word_acc", "nonword_acc", "wold_comp", "age_f8", "age_f9")

# List SDs in order above, for each group
stddev1 <- c(nara_comp_sd1, nara_acc_sd1, word_acc_sd1, nonword_acc_sd1, wold_comp_sd1, age_months_f8_sd, age_months_f9_sd)
stddev2 <- c(nara_comp_sd2, nara_acc_sd2, word_acc_sd2, nonword_acc_sd2, wold_comp_sd2, age_months_f8_sd, age_months_f9_sd)
stddev3 <- c(nara_comp_sd3, nara_acc_sd3, word_acc_sd3, nonword_acc_sd3, wold_comp_sd3, age_months_f8_sd, age_months_f9_sd)

# Create correlation matrix using above estimates
corr <- matrix(c(1, nara_comp_acc, nara_comp_word, nara_comp_nonword, nara_comp_wold, nara_comp_f8age, nara_comp_f9age,
                 nara_comp_acc, 1, nara_acc_word, nara_acc_nonword, nara_acc_wold, nara_acc_f8age, nara_acc_f9age,
                 nara_comp_word, nara_acc_word, 1, word_acc_nonword, word_acc_wold, word_acc_f8age, word_acc_f9age,
                 nara_comp_nonword, nara_acc_nonword, word_acc_nonword, 1, nonword_acc_wold, nonword_acc_f8age, nonword_acc_f9age,
                 nara_comp_wold, nara_acc_wold, word_acc_wold, nonword_acc_wold, 1, wold_f8age, wold_f9age,
                 nara_comp_f8age, nara_acc_f8age, word_acc_f8age, nonword_acc_f8age, wold_f8age, 1, f8age_f9age,
                 nara_comp_f9age, nara_acc_f9age, word_acc_f9age, nonword_acc_f9age, wold_f9age, f8age_f9age, 1),
               byrow = TRUE, nrow = 7, 
               dimnames = list(test_names, test_names))

# Compute covariance matrix, for each group
covar1 <- sdcor2cov(stddev1, corr)
covar2 <- sdcor2cov(stddev2, corr)
covar3 <- sdcor2cov(stddev3, corr)

################# SIMULATE DATA ###################

# List means in order above, for each group
means1 <- c(nara_comp_m1, nara_acc_m1, word_acc_m1, nonword_acc_m1, wold_comp_m1, age_months_f8_m, age_months_f9_m)
means2 <- c(nara_comp_m2, nara_acc_m2, word_acc_m2, nonword_acc_m2, wold_comp_m2, age_months_f8_m, age_months_f9_m)
means3 <- c(nara_comp_m3, nara_acc_m3, word_acc_m3, nonword_acc_m3, wold_comp_m3, age_months_f8_m, age_months_f9_m)

# Number of observations
n_ppts <- 1000

# Latent profile probabilities
class_probs<-c(0.6,0.15,0.25)

# Generate mixing profile indicator based on profile probabilities
k<-sample(1:3,n_ppts,class_probs,replace=TRUE)

# Set up empty holding dataframe
sim_dat <-data.frame(nara_comp=numeric(), nara_acc=numeric(), word_acc=numeric(), nonword_acc=numeric(), wold_comp=numeric(), age_f8=numeric(), age_f9 = numeric())

# Sample from the three multivariate normal distributions in the correct proportions according to profile probabilities
for(i in 1:n_ppts){
sim_dat[i,]<-switch(k[i],mvrnorm(n = 1, mu = means1, Sigma = covar1), mvrnorm(n = 1, mu = means2, Sigma = covar2), mvrnorm(n = 1, mu = means3, Sigma = covar3))
}
sim_dat <- as.data.frame(cbind(sim_dat,as.factor(k)))

# Create a dummy ID
sim_dat$id<-paste0("ALSPAC",sprintf('%0.4d', 1:n_ppts))

# Set variable names
names(sim_dat)<-c("naraComp", "naraAcc", "wordAcc", "nonwordAcc", "woldComp", "f8age", "f9age", "k","id")
```

### Missing data

**Missing data expectations for this analysis:**
* nara_comp and nara_acc should have no missing data, as inclusion in this/subsequent projects relies on having completed this assessment 
* word_acc & nonword_acc will also have only a very small proportion - collected in same assessment session as nara. Possibly more likely that if word_acc missing then nonword_acc also missing (?), but numbers very small. 
* wold_comp - more likely to have missing data; collected at different clinic session one year earlier 

```{r missing-data}
# specify variables collected in same session, low proportion missingness
same_na_cols <- c("wordAcc", "nonwordAcc")
same_na_prop <- 0.01

# specify variables collected in separate session, higher proportion missingness
sep_na_cols <- "woldComp"
sep_na_prop <- 0.05

# Create dataframe with NAs
sim_dat_NA <- sim_dat %>% 
  pivot_longer(cols = -c(id,k),names_to = "var", values_to = "value") %>%    # pivot data to long format
  mutate(r = runif(nrow(.)),                                            # simulate a random number (r) from 0 to 1 for each row
         value = ifelse(var %in% same_na_cols & r <= same_na_prop, NA,  # for same session vars, update to NA if r < threshold
                        ifelse(var %in% sep_na_cols & r <= sep_na_prop, NA, value))) %>%  # for separate session vars, update to NA is r < threshold, else same value
  dplyr::select(-r) %>%                                                        # remove random number
  pivot_wider(names_from = var, values_from = value)                    # pivot back to original format
```


### Data check

Inspect properties of the variables.

```{r summarise-vars, message = FALSE, warning = FALSE}
# Quick summary statistics
describe(sim_dat_NA)

# Overall histograms/correlations
sim_dat_NA %>% 
  dplyr::select(-k, -id) %>% 
  pairs.panels()

# Distributions/correlations by group
sim_dat_NA %>% 
  dplyr::select(-id) %>% 
  ggpairs(mapping = aes(color = k))

# Missing data
vis_miss(sim_dat_NA)
```

### Save data for use

Rename variables to match ALSPAC dataset.

```{r rename-round}
sim_dat_NA <- sim_dat_NA %>% 
  set_names(c("k", "yp_id", "nara_comp_raw_f9", "nara_acc_raw_f9", "read_word_raw_f9", "read_nonw_raw_f9", "wold_comp_raw_f8", "age_m_f8", "age_m_f9")) %>% 
  mutate_if(is.numeric, round) %>% 
  mutate(yp_no = as.numeric(as.factor(yp_id)))    # Create numeric version of ID for mplus
```

```{r save-simdata}
if(dir.exists("../data/simulated")==FALSE){dir.create("../data/simulated")} 

filename <- paste0("../data/simulated/WP1_SimpView_sim_raw_n", n_ppts, "_k", length(class_probs), ".csv")

write.csv(sim_dat_NA, filename, row.names = FALSE)
```
