---
title: "WP1: Simple View - Data Pre-processing and CFA"
output: 
  html_document:
    toc: true
    toc_float: true
date: '25/05/2021'
      
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output/analysis/") })
---

This script computes basic summary statistics for the data, ensures that they align with Mplus output, and tests the factor structure to be used for subsequent analyses. This analysis is conducted using the "exploratory" or "calibration" half of the dataset (subsample A).

# Set-up

```{r libraries, message = FALSE, warning = FALSE}
# List packages
pkgs <- c("tidyverse", "psych", "naniar", "MplusAutomation", "texreg", "semPlot", "knitr")

# Use groundhog package to load correct package versions
# if (!require("groundhog")) install.packages("groundhog")
# groundhog.day <- "2021-03-01"
# groundhog::groundhog.library(pkgs, groundhog.day)

# Or if not using groundhog package - load currently installed package versions separately (reproducibility not guaranteed)
invisible(lapply(pkgs, library, character.only = TRUE))
```

```{r create-dir}
# Create subdirectories for storing mplus scripts, data files, and output, if do not already exist
if(dir.exists("./mplus_models/")==FALSE){dir.create("./mplus_models/")}
if(dir.exists("./mplus_models/simpview/")==FALSE){dir.create("./mplus_models/simpview/")}
if(dir.exists("./mplus_models/simpview/desc")==FALSE){dir.create("./mplus_models/simpview/desc")}
if(dir.exists("./mplus_models/simpview/cfa")==FALSE){dir.create("./mplus_models/simpview/cfa")} 
```

###  Load and prepare data for Mplus

```{r load-data}
#sv_data <- read.csv("../data/simulated/WP1_SimpView_sim_raw_n1000_k3.csv") # for script testing 

sv_data <- read.csv("../data/processed/WP1_data_subA.csv")  %>%                # ALSPAC data
  dplyr::select(yp_id, cidB3153, age_m_f8, age_m_f9, 
                nara_acc_raw_f9, read_word_raw_f9, read_nonw_raw_f9,  # reading accuracy variables
                nara_comp_raw_f9, wold_comp_raw_f8,                   # comprehension variables
                ethn_white, m_age_birth, m_home_own, m_edu_alev)      # auxiliary variables
```

Mplus has an 8-character limit, so rename relevant variables with short names for modelling. 

```{r mplus-names}
# Check names
str_sub(names(sv_data), 1, 8)

# Rename any necessary, print new names
sv_data <- sv_data %>% 
  rename(naraComp = nara_comp_raw_f9,
         naraAcc = nara_acc_raw_f9,
         wordAcc = read_word_raw_f9,
         nonwAcc = read_nonw_raw_f9,
         woldComp = wold_comp_raw_f8,
         f8age = age_m_f8,
         f9age = age_m_f9) 

names(sv_data)
```

Variable format problematic for mplus data conversion, change ID to factor.

```{r mplus-formatting}
sv_data <- sv_data %>% 
  mutate(yp_id = as.factor(yp_id))
```


# Descriptive statistics

Extract summary statistics for each variable. Note that these will not actually reflect the final reported statistics, given that we are working here with only half of the sample. 

```{r summary-stats}
describe(sv_data) %>% 
  kable()

vis_miss(sv_data, cluster = TRUE)
```

Patterns of missingness were more thoroughly explored for all WP1 variables in the initial data processing script. The only variable with significant missingness (~15%) is the WOLD listening comprehension variable, as it was collected at a different clinic visit. We showed in the earlier script that the data met the assumption of missing at random, as missingness could be predicted by other variables in the dataset. We will therefore deal with missing data by using full information maximum likelihood. 

### Check that mplus reading data as expected

Extract descriptive statistics as check.

```{r mplus-desc}
# Specify model
m_desc <- mplusObject(
  TITLE = "Data check - Descriptive statistics;",
  ANALYSIS = "type = basic;",
  usevariables = c("naraComp", "naraAcc", "wordAcc", "nonwAcc", "woldComp", "f8age", "f9age"),
  rdata = sv_data)

# Fit model 
m_desc_fit <- mplusModeler(m_desc,
                            dataout = "./mplus_models/simpview/desc/sv_sim.dat",
                            modelout = "./mplus_models/simpview/desc/sv_check.inp",
                            check = TRUE, run = TRUE, hashfilename = TRUE)

# Read mplus output
m_desc_out <- readModels("./mplus_models/simpview/desc/sv_check.out")

# Check that sample sizes and means match descriptive statistics from above
r_summ <- describe(sv_data) %>%
  as.data.frame() %>% 
  slice(3:9) %>% 
  dplyr::select(n, mean) %>% 
  round(., 2) %>%
  rownames_to_column(var = "task") %>% 
  arrange(task)

mplus_summ <- m_desc_out$sampstat$univariate.sample.statistics %>% 
  as.data.frame() %>%
  round(.,2) %>% 
  dplyr::select(`Sample Size`, Mean) %>%
  rownames_to_column(var = "task") %>% 
  arrange(task)

if (any((r_summ[,2:3] - mplus_summ[,2:3]) > abs(0.01))){  # allowing for rounding errors
  print("WARNING: DIFFERENCES DETECTED BETWEEN R AND MPLUS DATA SUMMARIES")
  } else{
      print("Data check passed")
  }
```
(Only minor rounding errors)

# Confirmatory Factor Analysis

### One vs. Two Factors

Test whether the data support the proposed two-factor model (accuracy, comprehension) over a single reading factor.

Each model additionally includes age as a covariate. Variance of the age variables was included in the models explicitly so that participants with missing session data are not excluded in mplus.  

```{r cfa-factor-structure}
# ONE-FACTOR MODEL
m_cfa1 <- mplusObject(
  TITLE = "Confirmatory Factor Analysis - Single Factor;",
  ANALYSIS = "estimator = mlr; type = general;",                  
  MODEL = "f8age; f9age;
           read by naraComp naraAcc wordAcc nonwAcc woldComp;
           wordAcc nonwAcc naraAcc naraComp on f9age;
           woldcomp on f8age;",   # single reading factor, plus age covariates 
  OUTPUT = "sampstat; TECH1; TECH4; stdyx; modindices; ",
  PLOT = "TYPE = PLOT3;",
  usevariables = c("naraComp", "naraAcc", "wordAcc", "nonwAcc", "woldComp", "f8age", "f9age"),
  rdata = sv_data)

m_cfa1_fit <- mplusModeler(m_cfa1,
                           modelout = "./mplus_models/simpview/cfa/sv_cfa1.inp",
                           check = TRUE, run = FALSE)


# TWO-FACTOR MODEL
m_cfa2 <- update(m_cfa1,
                 TITLE = ~ "Confirmatory Factor Analysis - Two-Factor;",
                 MODEL = ~ "f8age; f9age;
                            acc by nonwAcc wordAcc naraAcc; comp by woldComp naraComp;
                            wordAcc nonwAcc naraAcc naraComp on f9age;
                            woldcomp on f8age;
                            acc with comp;
                            f8age with acc@0; f8age with comp@0;
                            f9age with acc@0; f9age with comp@0;",  # age-factor covariances fixed to 0 for identifiability
                 DEFINE = ~ "naraAcc = naraAcc/10; naraComp = naraComp/5;
                             f8age = f8age/2; f9age = f9age/2;") # divide by constant to make variances on similar scale


m_cfa2_fit <- mplusModeler(m_cfa2,
                           modelout = "./mplus_models/simpview/cfa/sv_cfa2.inp",
                           check = TRUE, run = FALSE)

# COMPARE MODELS
cfa_models <- readModels(target = "./mplus_models/simpview/cfa", filefilter = "sv_cfa")

# Inspect for model errors
cfa_models$sv_cfa1.out$warnings
cfa_models$sv_cfa2.out$warnings

#SummaryTable(cfa_models, keepCols = c("Title", "Parameters", "LL", "CFI", "TLI", "AIC", "BIC", "RMSEA_Estimate", "RMSEA_pLT05", "SRMR")) %>% 
#  kable()
```

A number of solutions were tried to resolve the issues with the two-factor model, including freeing the factor loadings, and rescaling the variables. However, the naracomp and woldcomp variables do not sit will together as a comprehension factor (and whereas naracomp sits better amongst the other reading variables).

On this basis, we will proceed with a Latent Profile Analysis. 


```{r cfa-inspection, include = FALSE}
# # Inspect model (by = loadings, with = covariances)
# cfa_models$sv_cfa2.out$parameters$unstandardized %>%
#   kable()
# cfa_models$sv_cfa2.out$parameters$stdyx.standardized %>%
#   kable()
# 
# # Inspect modification indices
# cfa_models$sv_cfa2.out$mod_indices %>%
#   arrange(desc(MI)) %>%
#   filter(operator == "WITH") %>%
#   kable()
```



```{r cfa-modification, include = FALSE}
# * Check for very large standardised residuals.  
# * No accepted view on how large MIs should be. Change only one at a time, starting with largest, but must be theoretically meaningful.  
# * *A priori*, we consider possible meaningful influences to be to be (in this order): naracomp with naraacc (taken from same assessment), wordacc with nonwacc (same battery of tests, similar format), wordacc with naraacc (both real word accuracy measures).  


# # TWO-FACTOR MODEL with additional covariance between the two single word reading tasks (word, nonword)
# m_cfa2b <- update(m_cfa2,
#                  TITLE = ~ "Confirmatory Factor Analysis - Two-Factor - modification1 single reading tasks;",
#                  MODEL = ~. + "nonwAcc with wordAcc;")
# 
# m_cfa2b_fit <- mplusModeler(m_cfa2b,
#                            modelout = "./mplus_models/simpview/cfa/sv_cfa2b.inp",
#                            check = TRUE, run = TRUE)
# 
# # COMPARE MODELS
# cfa_models <- readModels(target = "./mplus_models/simpview/cfa", filefilter = "sv_cfa")
# SummaryTable(cfa_models, keepCols = c("Title", "Parameters", "LL", "CFI", "TLI", "AIC", "BIC", "RMSEA_Estimate", "RMSEA_pLT05", "SRMR")) %>%
#   kable()


# The following rules of thumb are often used to indicate a good-fitting model:  
# 
# * CFI & TLI > .95  
# * RMSEA <= .05  
# * SRMR < .05
```


```{r final-cfa, include = FALSE}
# Inspect model (by = loadings, with = covariances)
# cfa_models$sv_cfa2b.out$parameters$unstandardized %>% 
#   kable()
# cfa_models$sv_cfa2b.out$parameters$stdyx.standardized %>% 
#   kable()
```

### CFA figure

```{r cfa-diagram}
semPaths(semPlotModel(cfa_models$sv_cfa1.out, mplusStd = "stdyx"), what = "paths", whatLabels = "std", rotation = 1, intercepts = FALSE)
```

### Auxiliary variables

```{r add-aux}
# Implement auxiliary variables

# Single-factor model with mplus implementation of auxiliary variables
m_final_aux <- update(m_cfa1,
                 TITLE = ~ "Confirmatory Factor Analysis - Final model with auxiliary variables",
                 VARIABLE = ~ . + "AUXILIARY = (M) ethn_white m_age_birth m_home_own m_edu_alev;",
                 usevariables = names(sv_data[!names(sv_data) %in% c("yp_id", "cidB3153")]),
                 rdata = sv_data)

m_final_aux_fit <- mplusModeler(m_final_aux,
                           modelout = "./mplus_models/simpview/cfa/sv_final_aux.inp",
                           check = TRUE, run = TRUE)


# Single-factor model with manual implementation of auxiliary variables (age 8 only)
m_aux_manual_part <- update(m_cfa1,
                 TITLE = ~ "Confirmatory Factor Analysis - Final model with auxiliary variables",
                 MODEL = ~ . + "f8age with ethn_white m_age_birth m_home_own m_edu_alev;
                                ethn_white with m_age_birth m_home_own m_edu_alev;
                                m_age_birth with m_home_own m_edu_alev;
                                m_home_own with m_edu_alev;
                                woldcomp with ethn_white m_age_birth 
                                m_home_own m_edu_alev;",
                 usevariables = names(sv_data[!names(sv_data) %in% c("yp_id", "cidB3153")]),
                 rdata = sv_data)

m_aux_manual_fit1 <- mplusModeler(m_aux_manual_part,
                           modelout = "./mplus_models/simpview/cfa/sv_final_aux_manualpart.inp",
                           check = TRUE, run = TRUE)

# Single-factor model with manual implementation of auxiliary variables (all)
m_aux_manual_full <- update(m_cfa1,
                 TITLE = ~ "Confirmatory Factor Analysis - Final model with auxiliary variables (full spec)",
                 MODEL = ~ . + "f8age with ethn_white m_age_birth m_home_own m_edu_alev;
                                f9age with ethn_white m_age_birth m_home_own m_edu_alev;
                 
                                ethn_white with m_age_birth m_home_own m_edu_alev;
                                m_age_birth with m_home_own m_edu_alev;
                                m_home_own with m_edu_alev;
                 
                                woldcomp with ethn_white m_age_birth 
                                m_home_own m_edu_alev;
                 
                                nonwacc with ethn_white m_age_birth 
                                m_home_own m_edu_alev;
                 
                                wordacc with ethn_white m_age_birth 
                                m_home_own m_edu_alev;
                 
                                naraacc with ethn_white m_age_birth 
                                m_home_own m_edu_alev;
                 
                                naracomp with ethn_white m_age_birth 
                                m_home_own m_edu_alev;
                 
                                woldcomp with ethn_white m_age_birth 
                                m_home_own m_edu_alev;",
                 usevariables = names(sv_data[!names(sv_data) %in% c("yp_id", "cidB3153")]),
                 rdata = sv_data)

m_aux_manual_fit2 <- mplusModeler(m_aux_manual_full,
                           modelout = "./mplus_models/simpview/cfa/sv_final_aux_manualfull.inp",
                           check = TRUE, run = TRUE)
```
