---
title: 'WP1: Simple View - Cross-validation'
output: 
  html_document:
    toc: true
    toc_float: true
    
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output/analysis/") })
---

This script is used for split-half cross validation. It takes the final model selected from WP1_SimpView_LPA_comp.Rmd (developed on sub-sample A) and tests it on sub-sample B, before re-running on the full dataset for reporting. The last section checks that this final model does not differ when auxiliary variables are included to support the MAR assumption.

The model running chunks are manually specified for the final model (E4) but the parameter-extraction is scripted, so requires (fairly minimal) editing if a different model is used.  

# Set-up 

```{r libraries}
# Specify packages
pkgs <- c("tidyverse", "MplusAutomation", "texreg", "kableExtra")

# Use groundhog package to load correct package versions
# if (!require("groundhog")) install.packages("groundhog")
# groundhog.day <- "2021-03-01"
# groundhog::groundhog.library(pkgs, groundhog.day)

# Or if not using groundhog package - load currently installed package versions separately (reproducibility not guaranteed)
invisible(lapply(pkgs, library, character.only = TRUE))
```

```{r create-dir}
# Create subdirectories for storing mplus scripts, data files, and output, if do not already exist
if(dir.exists("./mplus_models/")==FALSE){dir.create("./mplus_models/")}
if(dir.exists("./mplus_models/simpview/")==FALSE){dir.create("./mplus_models/simpview/")}
if(dir.exists("./mplus_models/simpview/lpa_comp")==FALSE){dir.create("./mplus_models/simpview/lpa_comp")} 
if(dir.exists("./mplus_models/simpview/lpa_comp/crossval")==FALSE){dir.create("./mplus_models/simpview/lpa_comp/crossval")}
if(dir.exists("../output/")==FALSE){dir.create("../output/")}
if(dir.exists("../output/figures")==FALSE){dir.create("../output/figures")}
if(dir.exists("../output/tables")==FALSE){dir.create("../output/tables")}
```

### Read in final model from exploratory sample

Read in selected model file that contains class assignment data, and data file of descriptive statistics. Merge and check. 

```{r read-model}
# Read in and extract class data
final_mod <- readModels(target = "./mplus_models/simpview/lpa_comp", filefilter = "e_rerun_4class")  # final model file
```

###  Load and prepare confirmatory subsample for Mplus

```{r load-data}
data_B <- read.csv("../data/processed/WP1_data_subB.csv")  %>%        # ALSPAC data sub B
  dplyr::select(yp_id, cidB3153, age_m_f8, age_m_f9, 
                nara_acc_raw_f9, read_word_raw_f9, read_nonw_raw_f9,  # reading accuracy variables
                nara_comp_raw_f9, wold_comp_raw_f8) %>%               # comprehension variables
  mutate(combAcc = read_word_raw_f9 + read_nonw_raw_f9) %>%           # create combined single item accuracy score
  select(-read_word_raw_f9, -read_nonw_raw_f9) %>%                    # remove other measures
  rename(naraComp = nara_comp_raw_f9,                                 # rename within mplus character limit
         naraAcc = nara_acc_raw_f9,
         woldComp = wold_comp_raw_f8,
         f8age = age_m_f8,
         f9age = age_m_f9) %>% 
  mutate(yp_id = as.factor(yp_id))
```

### Functions

Functions for processing the output are stored in a separate script, used for initial model fitting.
Also create new function for filling in Mplus model parameters with those extracted from exploratory model fitting. This is based on script from Grimm et al. (2017), downloaded from [first author's website](https://sites.google.com/site/longitudinalmethods/downloads?authuser=0)

```{r functions}
# Output Processing functions
source("WP1_OutputFunctions.R")

#loopReplace function is needed to fill in parameters in Mplus script with specified values
loopReplace <- function(text, replacements) {
	for (v in names(replacements)){
		text <- gsub(sprintf("\\[\\[%s\\]\\]", v), replacements[[v]], text)
	}
	return(text)
}
```

# Step 1: Evaluate model with fixed parameters

Based on script from Grimm et al. (2017), downloaded from [first author's website](https://sites.google.com/site/longitudinalmethods/downloads?authuser=0)

Extract model parameter estimates from final model selected in exploratory class enumeration process on subsample A (WP1_LPA_composite.Rmd)

```{r extract-subA-params}
params <- final_mod$parameters
df_params <- data.frame(params[[1]]$paramHeader,params[[1]]$param, params[[1]]$est, params[[1]]$LatentClass)
names(df_params) <- c("header", "param", "est", "class")

named_params <- df_params %>% 
  mutate(parameter_name = paste(header, param, class, sep = "_")) %>% 
  select(-c(header, param, class)) %>% 
  pivot_wider(names_from = parameter_name, values_from = est)
```

Create new model script fixing all parameters to estimated values above, and run with subsample B.
*Note that the model spec itself is written out manually, so will need editing if the final model set-up differs*

```{r run-fixed-model}
fixed_model <- mplusObject(
  TITLE = "LPA: Confirmatory model (subB fixed parameters)",
  ANALYSIS = "estimator = mlr; type = mixture; starts = 4000 1000; 
              processors = 4(STARTS);",
  VARIABLE = "idvariable = cidB3153; classes = c(4);",
  MODEL = loopReplace("
  %OVERALL% 
     f8age; f9age;
     combAcc naraAcc naraComp on f9age;
     woldcomp on f8age;
     naraComp; naraAcc; combAcc; woldComp;
     naraComp with naraAcc;
     naraComp with combAcc;
     naraComp with woldComp;
     naraAcc with combAcc;
     naraAcc with woldComp;
     combAcc with woldComp;
  
  [c#1@[[Means_C#1_Categorical.Latent.Variables]]];
  [c#2@[[Means_C#2_Categorical.Latent.Variables]]];
  [c#3@[[Means_C#3_Categorical.Latent.Variables]]];
  
  %c#1% 
     naraComp@[[Residual.Variances_NARACOMP_1]]; 
     naraAcc@[[Residual.Variances_NARAACC_1]]; 
     combAcc@[[Residual.Variances_COMBACC_1]];
     woldComp@[[Residual.Variances_WOLDCOMP_1]];
     naraComp with naraAcc@[[NARACOMP.WITH_NARAACC_1]];
     naraComp with combAcc@[[NARACOMP.WITH_COMBACC_1]];
     naraComp with woldComp@[[NARACOMP.WITH_WOLDCOMP_1]];
     naraAcc with combAcc@[[NARAACC.WITH_COMBACC_1]];
     naraAcc with woldComp@[[NARAACC.WITH_WOLDCOMP_1]];
     combAcc with woldComp@[[COMBACC.WITH_WOLDCOMP_1]];
  
  %c#2% 
     naraComp@[[Residual.Variances_NARACOMP_2]]; 
     naraAcc@[[Residual.Variances_NARAACC_2]]; 
     combAcc@[[Residual.Variances_COMBACC_2]];
     woldComp@[[Residual.Variances_WOLDCOMP_2]];
     naraComp with naraAcc@[[NARACOMP.WITH_NARAACC_2]];
     naraComp with combAcc@[[NARACOMP.WITH_COMBACC_2]];
     naraComp with woldComp@[[NARACOMP.WITH_WOLDCOMP_2]];
     naraAcc with combAcc@[[NARAACC.WITH_COMBACC_2]];
     naraAcc with woldComp@[[NARAACC.WITH_WOLDCOMP_2]];
     combAcc with woldComp@[[COMBACC.WITH_WOLDCOMP_2]];
  
  %c#3% 
     naraComp@[[Residual.Variances_NARACOMP_3]]; 
     naraAcc@[[Residual.Variances_NARAACC_3]]; 
     combAcc@[[Residual.Variances_COMBACC_3]];
     woldComp@[[Residual.Variances_WOLDCOMP_3]];
     naraComp with naraAcc@[[NARACOMP.WITH_NARAACC_3]];
     naraComp with combAcc@[[NARACOMP.WITH_COMBACC_3]];
     naraComp with woldComp@[[NARACOMP.WITH_WOLDCOMP_3]];
     naraAcc with combAcc@[[NARAACC.WITH_COMBACC_3]];
     naraAcc with woldComp@[[NARAACC.WITH_WOLDCOMP_3]];
     combAcc with woldComp@[[COMBACC.WITH_WOLDCOMP_3]];
  
  %c#4% 
     naraComp@[[Residual.Variances_NARACOMP_4]]; 
     naraAcc@[[Residual.Variances_NARAACC_4]]; 
     combAcc@[[Residual.Variances_COMBACC_4]];
     woldComp@[[Residual.Variances_WOLDCOMP_4]];
     naraComp with naraAcc@[[NARACOMP.WITH_NARAACC_4]];
     naraComp with combAcc@[[NARACOMP.WITH_COMBACC_4]];
     naraComp with woldComp@[[NARACOMP.WITH_WOLDCOMP_4]];
     naraAcc with combAcc@[[NARAACC.WITH_COMBACC_4]];
     naraAcc with woldComp@[[NARAACC.WITH_WOLDCOMP_4]];
     combAcc with woldComp@[[COMBACC.WITH_WOLDCOMP_4]];
  ", named_params), 
  OUTPUT = "TECH1 TECH8 TECH11 TECH14 stdyx CINTERVAL ENTROPY;",
  PLOT = "TYPE = PLOT3;",
  usevariables = colnames(data_B[,!names(data_B) %in% c("yp_id")]),
  rdata = data_B,
  SAVEDATA = "FILE IS fixed.dat; SAVE = cprobabilities;")

fixed_results <- mplusModeler(fixed_model, "mplus_models/simpview/lpa_comp/crossval/fixed.dat", run = FALSE)
```

Evaluate overall fit of model.
*except no good measures of absolute fit to use here? can only inspect general output and check for issues*

```{r fixed-out}
# Read in output
fixed_out <- readModels(target = "./mplus_models/simpview/lpa_comp/crossval", filefilter = "fixed")

# Print table 
fixed_out_summary <- lpa_enum_table(output = fixed_out)
fixed_out_summary

# Compute classification diagnostics
class_diag(fixed_out)

# Plot class means
plotMixtures_simpView(fixed_out)
```


# Step 2: Compare fit of fixed and freely estimated parameter models 

Fit the model to subsample B, allowing parameters to be freely estimated. 

Use previous parameters as starting values to guide the model into ordering the classes in the same way for comparison.
(*Not sure if I've done this correctly? Also could ignore this part if only interested in overall model fit*)

```{r free-model}
free_model <- mplusObject(
  TITLE = "LPA: Confirmatory model (subB free estimation)",
  ANALYSIS = "estimator = mlr; type = mixture; starts = 0; 
              ",
  VARIABLE = "idvariable = cidB3153; classes = c(4);",
  MODEL = loopReplace("
  %OVERALL% 
     f8age; f9age;
     combAcc naraAcc naraComp on f9age;
     woldcomp on f8age;
     naraComp; naraAcc; combAcc; woldComp;
     naraComp with naraAcc;
     naraComp with combAcc;
     naraComp with woldComp;
     naraAcc with combAcc;
     naraAcc with woldComp;
     combAcc with woldComp;
  
  [c#1*[[Means_C#1_Categorical.Latent.Variables]]];
  [c#2*[[Means_C#2_Categorical.Latent.Variables]]];
  [c#3*[[Means_C#3_Categorical.Latent.Variables]]];
  
  %c#1% 
     naraComp*[[Residual.Variances_NARACOMP_1]]; 
     naraAcc*[[Residual.Variances_NARAACC_1]]; 
     combAcc*[[Residual.Variances_COMBACC_1]];
     woldComp*[[Residual.Variances_WOLDCOMP_1]];
     naraComp with naraAcc*[[NARACOMP.WITH_NARAACC_1]];
     naraComp with combAcc*[[NARACOMP.WITH_COMBACC_1]];
     naraComp with woldComp*[[NARACOMP.WITH_WOLDCOMP_1]];
     naraAcc with combAcc*[[NARAACC.WITH_COMBACC_1]];
     naraAcc with woldComp*[[NARAACC.WITH_WOLDCOMP_1]];
     combAcc with woldComp*[[COMBACC.WITH_WOLDCOMP_1]];
  
  %c#2% 
     naraComp*[[Residual.Variances_NARACOMP_2]]; 
     naraAcc*[[Residual.Variances_NARAACC_2]]; 
     combAcc*[[Residual.Variances_COMBACC_2]];
     woldComp*[[Residual.Variances_WOLDCOMP_2]];
     naraComp with naraAcc*[[NARACOMP.WITH_NARAACC_2]];
     naraComp with combAcc*[[NARACOMP.WITH_COMBACC_2]];
     naraComp with woldComp*[[NARACOMP.WITH_WOLDCOMP_2]];
     naraAcc with combAcc*[[NARAACC.WITH_COMBACC_2]];
     naraAcc with woldComp*[[NARAACC.WITH_WOLDCOMP_2]];
     combAcc with woldComp*[[COMBACC.WITH_WOLDCOMP_2]];
  
  %c#3% 
     naraComp*[[Residual.Variances_NARACOMP_3]]; 
     naraAcc*[[Residual.Variances_NARAACC_3]]; 
     combAcc*[[Residual.Variances_COMBACC_3]];
     woldComp*[[Residual.Variances_WOLDCOMP_3]];
     naraComp with naraAcc*[[NARACOMP.WITH_NARAACC_3]];
     naraComp with combAcc*[[NARACOMP.WITH_COMBACC_3]];
     naraComp with woldComp*[[NARACOMP.WITH_WOLDCOMP_3]];
     naraAcc with combAcc*[[NARAACC.WITH_COMBACC_3]];
     naraAcc with woldComp*[[NARAACC.WITH_WOLDCOMP_3]];
     combAcc with woldComp*[[COMBACC.WITH_WOLDCOMP_3]];
  
  %c#4% 
     naraComp*[[Residual.Variances_NARACOMP_4]]; 
     naraAcc*[[Residual.Variances_NARAACC_4]]; 
     combAcc*[[Residual.Variances_COMBACC_4]];
     woldComp*[[Residual.Variances_WOLDCOMP_4]];
     naraComp with naraAcc*[[NARACOMP.WITH_NARAACC_4]];
     naraComp with combAcc*[[NARACOMP.WITH_COMBACC_4]];
     naraComp with woldComp*[[NARACOMP.WITH_WOLDCOMP_4]];
     naraAcc with combAcc*[[NARAACC.WITH_COMBACC_4]];
     naraAcc with woldComp*[[NARAACC.WITH_WOLDCOMP_4]];
     combAcc with woldComp*[[COMBACC.WITH_WOLDCOMP_4]];", named_params), 
  OUTPUT = "TECH1 TECH8 TECH11 TECH14 stdyx CINTERVAL ENTROPY;",
  PLOT = "TYPE = PLOT3;",
  usevariables = colnames(data_B[,!names(data_B) %in% c("yp_id")]),
  rdata = data_B,
  SAVEDATA = "FILE IS free.dat; SAVE = cprobabilities;")

free_results <- mplusModeler(free_model, "mplus_models/simpview/lpa_comp/crossval/free.dat", run = TRUE)
```

```{r compare-fit}
# Use readModels to create extract all info for both models
all_mods <- readModels("./mplus_models/simpview/lpa_comp/crossval")

compareModels(all_mods[["fixed.out"]], all_mods[["free.out"]],
              show = c("diff", "pdiff", "summaries", "unique"),
              equalityMargin = c(param = .05, pvalue = .02),
              sort = "type", diffTest = TRUE, showNS = FALSE)

```

If difference in fit, should probably repeat class enumeration process with subset B, see if arrive at same model? 

*****************************************************************************************
TO HERE - remaining script not run/tested yet 
*****************************************************************************************

# Step 3: Re-fit the model on the full dataset for reporting 

```{r full-data}
data_all <- read.csv("../data/processed/WP1_data_all.csv")  %>%       # ALSPAC data
  dplyr::select(yp_id, cidB3153, age_m_f8, age_m_f9, 
                nara_acc_raw_f9, read_word_raw_f9, read_nonw_raw_f9,  # reading accuracy variables
                nara_comp_raw_f9, wold_comp_raw_f8,                   # comprehension variables
                ethn_white, m_age_birth, m_home_own, m_edu_alev) %>%  # auxiliary variables
  rename(naraComp = nara_comp_raw_f9,
         naraAcc = nara_acc_raw_f9,
         wordAcc = read_word_raw_f9,
         nonwAcc = read_nonw_raw_f9,
         woldComp = wold_comp_raw_f8,
         f8age = age_m_f8,
         f9age = age_m_f9,
         ethn = ethn_white,
         mAge = m_age_birth,
         home = m_home_own,
         mEdu = m_edu_alev) %>% 
  mutate(yp_id = as.factor(yp_id))
```

```{r full-model}
full_model <- mplusObject(
  TITLE = "Latent Profile Analysis; Final model full dataset",
  ANALYSIS = "estimator = mlr; type = mixture; starts = 4000 1000; 
              processors = 4(STARTS);",
  VARIABLE = "idvariable = cidB3153; classes = c(4);",
  MODEL = "
  %OVERALL% 
     f8age; f9age;
     combAcc naraAcc naraComp on f9age;
     woldcomp on f8age;
     naraComp; naraAcc; combAcc; woldComp;
     naraComp with naraAcc;
     naraComp with combAcc;
     naraComp with woldComp;
     naraAcc with combAcc;
     naraAcc with woldComp;
     combAcc with woldComp;
  
  %c#1% 
     naraComp; 
     naraAcc; 
     combAcc;
     woldComp;
     naraComp with naraAcc;
     naraComp with combAcc;
     naraComp with woldComp;
     naraAcc with combAcc;
     naraAcc with woldComp;
     combAcc with woldComp;
  
  %c#2% 
     naraComp; 
     naraAcc; 
     combAcc;
     woldComp;
     naraComp with naraAcc;
     naraComp with combAcc;
     naraComp with woldComp;
     naraAcc with combAcc;
     naraAcc with woldComp;
     combAcc with woldComp;
  
  %c#3% 
     naraComp; 
     naraAcc; 
     combAcc;
     woldComp;
     naraComp with naraAcc;
     naraComp with combAcc;
     naraComp with woldComp;
     naraAcc with combAcc;
     naraAcc with woldComp;
     combAcc with woldComp;
  
  %c#4% 
     naraComp; 
     naraAcc; 
     combAcc;
     woldComp;
     naraComp with naraAcc;
     naraComp with combAcc;
     naraComp with woldComp;
     naraAcc with combAcc;
     naraAcc with woldComp;
     combAcc with woldComp;", 
  OUTPUT = "TECH1 TECH8 TECH11 TECH14 stdyx CINTERVAL ENTROPY;",
  PLOT = "TYPE = PLOT3;",
  usevariables = colnames(sv_data[,!names(sv_data) %in% c("yp_id", "ethn", "mAge", "home", "mEdu")]),
  rdata = data_all,
  SAVEDATA = "FILE IS free.dat; SAVE = cprobabilities;")

full_results <- mplusModeler(free_model, "mplus_models/simpview/lpa_comp/crossval/full.dat", run = TRUE)

```

# Check that auxiliary variables do not influence results

```{r auxiliary}
# Single-factor model with manual implementation of auxiliary variables (all)
m_aux_manual_full <- update(full_model,
                 TITLE = ~ "Confirmatory Factor Analysis - Final model with auxiliary variables (full spec)",
                 MODEL = ~ . + "f8age with ethn mAge home mEdu;
                                f9age with ethn mAge home mEdu;
                 
                                ethn with mAge home mEdu;
                                mAge with home mEdu;
                                home with mEdu;
                 
                                woldcomp with ethn mAge 
                                home mEdu;
                 
                                combacc with ethn mAge 
                                home mEdu;
                 
                                naraacc with ethn mAge 
                                home mEdu;
                 
                                naracomp with ethn mAge 
                                home mEdu;
                 
                                ",
                 usevariables = names(sv_data[!names(sv_data) %in% c("yp_id", "cidB3153")]),
                 rdata = data_full)

aux_results <- mplusModeler(free_model, "mplus_models/simpview/lpa_comp/crossval/aux.dat", run = TRUE)


# Compare fit ???? Or just look at general interpretability?? 
all_mods <- readModels("../mplus_models/simpview/lpa_comp/crossval/")

compareModels(all_mods[["full.out"]], all_mods[["aux.out"]],
              show = c("diff", "pdiff", "summaries", "unique"),
              equalityMargin = c(param = .05, pvalue = .02),
              sort = "type", diffTest = TRUE, showNS = FALSE)
```
